\documentclass{article}
\usepackage[left=3cm,right=3cm,top=3cm,bottom=3cm]{geometry}
\usepackage{amsmath,amssymb,amsthm,pgfplots,tikz}
\usepackage[inline]{enumitem}
\usepackage{color}
\setlength{\parindent}{0mm} %So that we do not indent on new paragraphs
\newcommand{\TODO}[1]{\textcolor{red}{TODO: #1}}

\begin{document}
\title{Graduate Algebra II: Homework 5}
\author{Li Ling Ko\\ lko@nd.edu}
\date{\today}
\maketitle

\it \textbf{Section 12.2 Q3:} Prove that two $2\times2$ matrices over $F$
  which are not scalar matrices are similar if and only if they have the
  same characteristic polynomial.

  \begin{proof}
    We first show that similar matrices have the same characteristic
    polynomial. Given similar matrices $A$ and $P^{-1}AP$,
    \begin{align*}
      \text{char}(P^{-1}AP) &=|xI-P^{-1}AP|\\
      &=|P(xI-P^{-1}AP)P^{-1}|\\
      &=|P(xI)P^{-1}-A|\\
      &=|xI-A|\\
      &=\text{char}(A).\\
    \end{align*}

    For the converse, let $A,B\in\text{Mat}_2(F)$ be two non-scalar
    matrices over with the same characteristic polynomial. Suppose the
    characteristic polynomial of $A$ and $B$ has two distinct roots, or is
    an irreducible polynomial of degree two. Then the minimal polynomial of
    the matrices must also be the characteristic polynomial by Proposition
    20, so the matrices have the same rational canonical form, and
    are therefore similar to each other from Theorem 17. \\

    On the other hand, if the roots of the characteristic polynomial are
    not distinct, i.e.
    \[c_A(x)=c_B(x)=(x-\alpha)^2\]
    for some $\alpha\in F$, then $m_A(x)$ is either $x-\alpha$ or
    $(x-\alpha)^2$. If $m_A(x)=x-\alpha$, then the invariant factors of $A$
    are $x-\alpha$ and $x-\alpha$, which implies that the rational
    canonical form of $A$ is $\alpha I$. Now since $A$ is similar to its
    rational canonical form $\alpha I$,we have $A=Q^{-1}(\alpha I)Q=\alpha
    I$ for some invertible matrix $Q\in\text{Mat}_2(F)$, which makes $A$ a
    scalar matrix, a contradiction. Therefore we must have
    $m_A(x)=m_B(x)=(x-\alpha)^2$, and so $A$ and $B$ have the same rational
    canonical form, making them similar by Theorem 17.
  \end{proof}

\it \textbf{Section 12.2 Q4:} Prove that two $3\times3$ matrices are
  similar if and only if they have the same characteristic and same minimal
  polynomial. Give an explicit counterexample to this assertion for
  $4\times4$ matrices.

  \begin{proof}
    Assume $A,B\in\text{Mat}_3(F)$ are similar. Then from the argument in
    the previous question, the $A$ and $B$ have the same characteristic
    polynomial. Also, from Theorem 15, $A$ and $B$ have the same rational
    canonical form and thus also have the same minimal polynomial.\\

    For the converse, assume $A,B\in\text{Mat}_3(F)$ have the same
    characteristic polynomial $c(x)$ and minimal polynomial $m(x)$. If
    $c(x)$ has distinct roots, or is an irreducible polynomial of degree 3,
    or is the product of an irreducible polynomial of degree 2 with a
    linear factor, then by Proposition 20, $A$ and $B$ will have the same
    invariant factors, and therefore are similar from Theorem 17. \\

    On the other hand, if $c(x)$ has only two distinct roots, i.e.
    \[c_A(x)=c_B(x)=c(x) =(x-\alpha)(x-\beta)^2\]
    for some $\alpha\neq\beta\in F$, the minimal polynomial $m(x)$ of $A$
    and $B$ can either be $c(x)$ or $(x-\alpha)(x-\beta)$. If
    $m(x)=c(x)$, then the invariant factors of $A$ and of $B$ can only be
    $m(x)$, and then $A$ would be similar to $B$ from Theorem 17. If
    $m(x)=(x-\alpha)(x-\beta)$, then the invariant factors of both matrices
    can only be $(x-\beta)$ and $(x-\alpha)(x-\beta)$. Then again, $A$ and
    $B$ are similar from Theorem 17. \\

    Finally, if $c(x)$ has only one root, i.e.
    \[c_A(x)=c_B(x)=c(x) =(x-\alpha)^3\]
    for some $\alpha\in F$, the minimal polynomial $m(x)$ of $A$ and $B$
    can either be $x-\alpha$, $(x-\alpha)^2$, or $(x-\alpha)^3$. If it is
    the first case, then the invariant factors of $A$ and of $B$ can only
    be $x-\alpha$ three times, and then $A$ would be similar to $B$ from
    Theorem 17. If it is the second case, then the invariant factors can
    only be $x-\alpha$ and $(x-\alpha)^2$, and again $A$ would be similar
    to $B$. Finally, if it is the third case, then the invariant factors
    can only be $(x-\alpha)^3$, and then $A$ and $B$ would still be
    similar from Theorem 17. \\

    Thus two $3\times3$ matrices are similar if and only if they have the
    same characteristic and minimal polynomial. The assertion for
    $4\times4$ matrices $A$ and $B$ will fail if for both matrices, the
    characteristic polynomial is $(x-\alpha)^4$ and the minimal polynomial
    is $(x-\alpha)^2$. Then if $A$ has invariant factors $(x-\alpha)^2$ and
    $(x-\alpha)^2$ and $B$ has invariant factors $x-\alpha$, $x-\alpha$,
    and $(x-\alpha)^2$, by Theorem 17, their respective rational canonical
    forms $A'$ and $B'$ will not be similar even though they have the same
    characteristic and minimal polynomials.
  \end{proof}

\it \textbf{Section 12.2 Q6:} Prove that the constant term in the
  characteristic polynomial of the $n\times n$ matrix $A$ is
  $(-1)^n\text{det}A$ and that the coefficient of $x^{n-1}$ is the negative
  of the sum of the diagonal entries of $A$ (the sum of the diagonal
  entries of $A$ is called the trace of $A$). Prove that $\text{det}A$ is
  the product of the eigenvalues of $A$ and that the trace of $A$ is the
  sum of the eigenvalues of $A$.

  \begin{proof}
    If the eigenvalues do not lie in the field, we can work in the
    splitting field extension of the minimal polynomial. Therefore we can
    assume that the eigenvalues exist in the field we are working in. The
    constant term of $c_A(x)$ is
    \[c_A(0)=\text{det}(0I-A)=\text{det}(-A)=(-1)^n\text{det}(A).\]

    To prove the assertion on the coefficient of $x^{n-1}$, first note that
    given two matrices $X$ and $Y$, we have $\text{tr}(XY)=\text{tr}(YX)$,
    and therefore similar matrices will have the same trace. Hence it
    suffices to prove that the coefficient of $x^{n-1}$ is the same as the
    trace of the rational canonical form of $A$. By definition, if the
    invariant factors of $A$ are $m_1(x),\ldots,m_k(x)$, then the trace of
    the rational canonical form of $A$ is $-\sum_{i=1}^k c_{i}$, where
    $c_{i}$ is the coefficient of the second highest degree of $m_i(x)$.
    Now since similar matrices have the same characteristic polynomial, we
    can assume that the matrix $A$ is of rational canonical form. Then the
    coefficient of $x^{n-1}$ of $c_A(x)=m_1(x)\cdots m_k(x)$ will be
    $-\sum_{i=1}^k c_{i}$, as required. \\

    The $n$ eigenvalues of $A$ (including multiplicities)
    $\lambda_1,\ldots,\lambda_n$ are exactly the set of
    roots of the characteristic polynomial of $A$, i.e.
    \[c_A(x) =(x-\lambda_1)\cdots(x-\lambda_n).\]
    Setting $x=0$ in the above equation gives us
    $c_A(0)=(-1)^n\lambda_1\cdots\lambda_n$. And since $c_A(0)$ is
    $(-1)^n\text{det}(A)$ as shown earlier, we have
    \[\text{det}(A) =\lambda_1\cdots\lambda_n.\]

    Similarly, in the expansion of $c_A(x)$, the coefficient of $x^{n-1}$
    is the negative of the sum of the eigenvalues of $A$. And since we have
    shown earlier that the former is the same as $-\text{tr}(A)$, we get
    that the sum of the eigenvalues is the same as $\text{tr}(A)$.
  \end{proof}

\it \textbf{Section 12.2 Q7:} Determine the eigenvalues of the matrix
  \[A=\begin{pmatrix}
    0&1&0&0\\
    0&0&1&0\\
    0&0&0&1\\
    1&0&0&0\\
  \end{pmatrix}.\]

  \begin{proof}
    Observe that $A$ describes a permutation of order four, and therefore
    satisfies $A^4=I$. So $m_A(x)$ must divide $x^4-1$, which has distinct
    roots $1,-1,i,-i$. Since $c_A(x)$ has the same roots as $m_A(x)$ and
    the roots of $c_A(x)$ are the eigenvalues of $A$, therefore the
    eigenvalues of $A$ can only be from $\{1,-1,i,-i\}$. We observe that
    $1$ is an eigenvalue via eigenvector $[1,1,1,1]^T$. Thus, from
    Question 6, for the eigenvalues to sum to 0, -1 must also be an
    eigenvalue. Furthermore, since $\text{det}(iI-A)=0$, $i$ is also an
    eigenvalue, and so for the eigenvalues to sum to 0 again, $-i$ is also
    an eigenvalue. Thus the four eigenvalues are \[1,-1,i,-i.\]
  \end{proof}

\it \textbf{Section 12.2 Q9:} Find the rational canonical forms of
  \begin{align*}
    A=\begin{pmatrix}
      0&-1&-1\\
      0&0&0\\
      -1&0&0\\
    \end{pmatrix}, \text{and}\;
    B=\begin{pmatrix}
      c&0&-1\\
      0&c&1\\
      -1&1&c\\
    \end{pmatrix}.
  \end{align*}

  \begin{proof}
    We first find the eigenvalues of $A$. Since the products of the
    eigenvalues must equal $(-1)^3\text{det}(A)=0$, one of the eigenvalues
    must be 0. Also, the sum of the eigenvalues must equal the trace of $A$
    which is 0, so the remaining two eigenvalues must be additive inverses.
    Observing that $[1,0,1]^T$ is an eigenvector with eigenvalue -1, we get
    that the eigenvalues must be $\{0,1,-1\}$. Then from Proposition 20,
    since the eigenvalues are distinct, $A$ can only have one invariant
    factor, which is $m_A(x)=x(x-1)(x+1)=x^3-x$. Thus the rational
    canonical form of $A$ is the single companion matrix
    $\mathcal{C}_{m_A(x)}$, given by
    \[\begin{pmatrix}
      0&0&0\\
      1&0&1\\
      0&1&0\\
    \end{pmatrix}.\]

    Next, we find the eigenvalues of $C$. Note that $B=cI+C$, where
    \begin{align*}
      C=\begin{pmatrix}
        0&0&-1\\
        0&0&1\\
        -1&1&0\\
      \end{pmatrix}.
    \end{align*}

    Therefore
    \begin{align*}
      c_B(x) &=\text{det}(xI-B)\\
      &=\text{det}(xI-cI-C)\\
      &=\text{det}((x-c)I-C)\\
      &=c_C(x-c).
    \end{align*}

    Since the eigenvalues of $C$ are the roots of $c_C(x)$, the eigenvalues
    of $B$ must be the same roots but plus $c$. Thus we attempt to find the
    eigenvalues of $C$. We have shown in Question 6 the sum of the
    eigenvalues is the trace of $C$, which is 0, and their product is
    $(-1)^3\text{det}(C)=0$. Thus one of the eigenvalues must be 0, and the
    other two eigenvalues must be additive inverses. Indeed,
    \[c_C(x)=\text{det}(xI-C)=x^3-2x,\]
    which has distinct roots $\{0,\sqrt{2},-\sqrt{2}\}$. \\

    Then
    \[c_B(x)=c_C(x-c)=(x-c)^3-2(x-c) =x^3-3cx^2+(3c^2-2)x+(2c-c^3),\]
    and the eigenvalues of $B$ must be distinct since those of $C$ are also
    distinct. Thus from Proposition 20, $B$ can have only one invariant
    factor, which is the same as its characteristic polynomial. Hence the
    rational canonical form of $B$ is the single companion matrix
    $\mathcal{C}_{c_B(x)}$, given by
    \[\begin{pmatrix}
      0&0&c^3-2c\\
      1&0&2-3c^2\\
      0&1&3c\\
    \end{pmatrix}.\]
  \end{proof}

\it \textbf{Section 12.2 Q10:} Find all similarlity classes of $6\times6$
  matrices over $\mathbb{Q}$ with minimal polynomial $m(x)=(x+2)^2(x-1)$ (it
  suffices to give all lists of invariant factors and write out some of
  their corresponding matrices).

  \begin{proof}
    From Proposition 20, the remaining invariant factors can only have
    roots -2 or 1, and must divide $m(x)$. Furthermore, the products the
    invariant factors must have degree 6, and the first invariant factor
    must divide the second and so on. Finally, the similarity classes is
    completely characterized by its corresponding set of invariant factors.
    Thus Table~\ref{tbl:classes} are the similarity classes, with some
    respective rational canonical forms:

    \begin{table}[h]
    \centering
    \begin{tabular}{lc}
      Invariant factors &Rational canonical form\\\hline\\
      $(x+2)^2(x-1)$, $(x+2)^2(x-1)$
        &$\begin{pmatrix}
          0&0&4&0&0&0\\
          1&0&0&0&0&0\\
          0&1&-3&0&0&0\\
          0&0&0&0&0&4\\
          0&0&0&1&0&0\\
          0&0&0&0&1&-3\\
        \end{pmatrix}$\\\\
      $x+2$, $(x+2)(x-1)$, $(x+2)^2(x-1)$
        &$\begin{pmatrix}
          -2&0&0&0&0&0\\
          0&0&2&0&0&0\\
          0&1&-1&0&0&0\\
          0&0&0&0&0&4\\
          0&0&0&1&0&0\\
          0&0&0&0&1&-3\\
        \end{pmatrix}$\\\\
      $x-1$, $(x+2)(x-1)$, $(x+2)^2(x-1)$
        &$\begin{pmatrix}
          1&0&0&0&0&0\\
          0&0&2&0&0&0\\
          0&1&-1&0&0&0\\
          0&0&0&0&0&4\\
          0&0&0&1&0&0\\
          0&0&0&0&1&-3\\
        \end{pmatrix}$\\\\
      $x+2$, $(x+2)^2$, $(x+2)^2(x-1)$&\\
      $x+2$, $x+2$, $x+2$, $(x+2)^2(x-1)$&\\
      $x-1$, $x-1$, $x-1$, $(x+2)^2(x-1)$&\\
    \end{tabular}
    \caption{Similarity classes where $m(x)=(x+2)^2(x-1)$}
    \label{tbl:classes}
    \end{table}
  \end{proof}

\it \textbf{Section 12.2 Q11:} Find all similarity classes of $6\times6$
  matrices over $\mathbb{C}$ with characteristic polynomial
  $c(x)=(x^4-1)(x^2-1)$.

  \begin{proof}
    Repeating the argument used in the previous question, since
    \[c(x)=(x^4-1)(x^2-1)=(x+i)(x-i)(x+1)^2(x-1)^2,\]
    the only possible similarity classes as characterized by the set of
    invariant factors are:

    \begin{enumerate}
      \item $(x+i)(x-i)(x+1)^2(x-1)^2$
      \item $(x+1)$, $(x+i)(x-i)(x+1)(x-1)^2$
      \item $(x-1)$, $(x+i)(x-i)(x+1)^2(x-1)$
      \item $(x+1)(x-1)$, $(x+i)(x-i)(x+1)(x-1)$
    \end{enumerate}
  \end{proof}

\it \textbf{Section 12.2 Q12:} Find all similarity classes of $3\times3$
  matrices $A$ over $\mathbb{F}_2$ satisfying $A^6=I$ (compare with the
  answer we computed over $\mathbb{Q}$). Do the same for $4\times4$ matrices
  $B$ satisfying $B^{20}=I$.

  \begin{proof}
    $x^6-1$ factorizes into:
    \[x^6-1=(x-1)(x+1)(x^2+x+1)(x^2-x+1).\]

    Over $\mathbb{F}_2$, since the quadratic factors do not have roots,
    they must be irreducible. Then from irreducibility criteria
    (Proposition 12 Section 9.4), the quadratic factors are also
    irreducible over $\mathbb{Z}$, which makes them irreducible over
    $\mathbb{Q}$ (Proposition 11 Section 9.4). Therefore the factorization
    given above gives factors that are irreducible over $\mathbb{F}_2$ and
    $\mathbb{Q}$. \\

    Now $m_A(x)$ is the monic polynomial of smallest degree that
    annihilates $A$, therefore it must divide $x^6-1$. Furthermore,
    $c_A(x)$ must have the same roots as $m_A(x)$, must be a multiple of
    $m_A(x)$, and must have degree 3, therefore the only possible pairs of
    minimal and characteristic polynomials are as shown in
    Table~\ref{tbl:A1} for $\mathbb{Q}$ and in Table~\ref{tbl:A2} for
    $\mathbb{F}_2$. Now since similarity classes of $3\times3$ matrices are
    characterized by the characteristic and minimal polynomial pairs
    (Question 6), the pairings in Tables~\ref{tbl:A1} and~\ref{tbl:A2} are
    exactly the similarity classes for both $\mathbb{Q}$ and $\mathbb{F}_2$
    respectively. \\

    \begin{table}[h]
    \centering
    \begin{tabular}{ll}
      Minimal polynomial &Characteristic polynomial\\\hline\\
      $(x-1)(x^2+x+1)$ &$(x-1)(x^2+x+1)$\\
      $(x+1)(x^2+x+1)$ &$(x+1)(x^2+x+1)$\\
      $(x-1)(x^2-x+1)$ &$(x-1)(x^2-x+1)$\\
      $(x+1)(x^2-x+1)$ &$(x+1)(x^2-x+1)$\\
      $(x-1)(x+1)$ &$(x-1)(x+1)^2$\\
      $(x-1)(x+1)$ &$(x-1)^2(x+1)$\\
      $(x-1)$ &$(x-1)^3$\\
      $(x+1)$ &$(x+1)^3$\\
    \end{tabular}
    \caption{All possible pairs of characteristic and minimal polynomials
      over $\mathbb{Q}$}
    \label{tbl:A1}
    \end{table}

    \begin{table}[h]
    \centering
    \begin{tabular}{ll}
      Minimal polynomial &Characteristic polynomial\\\hline\\
      $(x-1)(x^2+x+1)$ &$(x-1)(x^2+x+1)$\\
      $(x-1)^2$ &$(x-1)^3$\\
      $(x-1)$ &$(x-1)^3$\\
    \end{tabular}
    \caption{All possible pairs of characteristic and minimal polynomials
      over $\mathbb{F}_2$}
    \label{tbl:A2}
    \end{table}

    For $4\times4$ matrices $B$ satisfying $B^{20}=1$, we first factorize
    $x^{20}-1$ over $\mathbb{F}_2$ as follows:
    \begin{align*}
      x^{20}-1 &=(x^{10}-1)(x^{10}+1)\\
      &=(x^{10}-1)^2\\
      &=((x^{5}-1)(x^{5}+1))^2\\
      &=(x^{5}-1)^4\\
      &=((x-1)(x^4+x^3+x^2+x+1))^4\\
      &=(x-1)^4(x^4+x^3+x^2+x+1)^4.\\
    \end{align*}

    We show that the quartic factor is irreducible in $\mathbb{F}_2$: It
    has no linear factors since it it has no roots in $\mathbb{F}_2$. Thus
    it can only have quadratic factors. Then we will be able to write 
    \[(x^4+x^3+x^2+x+1) =(x^2+ax+1)(x^2+bx+1).\]
    But comparing coefficients for $x^2$ gives us $a=b=1$, yet comparing
    coefficients for $x^3$ gives us $a+b=1$, a contradiction. \\

    Therefore the factorization given above gives factors that are
    irreducible over $\mathbb{F}_2$. Then repeating the argument above to
    get the minimal polynomial, the only possibilities for the minimal
    polynomial are
    \[m_B(x)=x-1, (x-1)^2, (x-1)^3, (x-1)^4,\;\; \text{or}\;\;
    x^4+x^3+x^2+x+1.\]

    For each of these minimal polynomials, the possible invariant factors
    that will characterize the similarity classes are as given by:
    \begin{enumerate}
      \item $x^4+x^3+x^2+x+1$
      \item $(x-1)^4$
      \item $x-1$, $(x-1)^3$
      \item $(x-1)^2$, $(x-1)^2$
      \item $x-1$, $x-1$, $(x-1)^2$
      \item $x-1$, $x-1$, $x-1$, $x-1$.
    \end{enumerate}
  \end{proof}

\it \textbf{Section 12.2 Q13:} Prove that the number of similarity
  classes of $3\times3$ matrices over $\mathbb{Q}$ with a given
  characteristic polynomial in $\mathbb{Q}[x]$ is the same as the number of
  similarity classes over any extension field of $\mathbb{Q}$. Give an
  example to show that this is not true in general for $4\times4$ matrices.

  \begin{proof}
    Observe that as the extension field gets larger, the characteristic
    polynomial $c(x)$ can be split into smaller factors, and therefore will
    have more similarity classes, because there will be more possibilities
    for sets of invariant factors, and also because invariant factors from
    the smaller field remain valid invariant factors over the larger
    field.  Also observe that $\mathbb{C}$, the completion of
    $\mathbb{Q}$, will give the maximum number of similarity classes
    since the characteristic polynomial will be factorized completely
    into linear factors. Thus it suffices to show that the number of
    similarity classes over $\mathbb{Q}$ is the same as the number of
    similarity classes over $\mathbb{C}$. \\

    Let $c(x)\in\mathbb{Q}[x]$ be the characteristic polynomial of degree
    three. If $c(x)$ has distinct roots in $\mathbb{C}$, then the minimal
    polynomial can only be $c(x)$ itself, giving a single invariant factor
    equal to $c(x)$ and therefore a single similarity class. Also, if all
    the roots of $c(x)$ lie in $\mathbb{Q}$, then $c(x)$ will factorize
    linearly over $\mathbb{Q}$, and so the possible sets of invariant
    factors will be the same over $\mathbb{Q}$ or over $\mathbb{C}$. Thus
    we can assume that $c(x)$ has at most two distinct roots, and at least
    one of them lies outside $\mathbb{Q}$. \\

    Assume $c(x)=(x-a)(x-b)^2\in\mathbb{Q}[x]$ has two distinct roots, at
    least one of which lies outside $\mathbb{Q}$. Then the minimal
    polynomial $m(x)\in\mathbb{Q}[x]$ can only be either $(x-a)(x-b)$ or
    $c(x)$. If it is the former, then the coefficients of $c(x)$
    give $a+2b\in\mathbb{Q}$, and the coefficeints of $m(x)$ give
    $a+b\in\mathbb{Q}$. Thus, $b\in\mathbb{Q}$, and therefore also
    $a\in\mathbb{Q}$, which would contradict $a,b\not\in\mathbb{Q}$. Hence
    the minimal polynomial can only be $c(x)$, giving only one possible
    similarity class, which is the one associated with the set of invariant
    factors containing only $c(x)$ itself. \\

    Finally, assume $c(x)=(x-a)^3\in\mathbb{Q}[x]$ has only one root $a$
    which lies outside $\mathbb{Q}$. But the coefficient of $x^2$ gives
    $3a\in\mathbb{Q}$, which implies $a\in\mathbb{Q}$, a contradiction.
  \end{proof}

\it \textbf{Section 12.2 Q19:} Let $V$ be the infinite dimensional real
  vector space
  \[\mathbb{R}^\infty =\{(a_0,a_1,a_2,\ldots)|a_0,a_1,a_2\in\mathbb{R}.\}\]
  Define the map $T:V\rightarrow V$ by
  $T(a_0,a_1,a_2\ldots)=(0,a_0,a_1,a_2,\ldots)$. Prove that $T$ has no
  eigenvectors.

  \begin{proof}
    Assume that $T$ has eigenvector $v=(a_0,a_1,\ldots)\neq0$. Then
    \begin{align*}
      (0,a_0,a_1,\ldots) &=Tv\\
      &=\lambda v\\
      &=(\lambda a_0,\lambda a_1,\ldots).\\
    \end{align*}
    for some $\lambda\in\mathbb{R}$. Therefore $\lambda a_0=0$, and
    $\lambda a_{i+1}=a_i$ for all $i\in\mathbb{N}$. If $\lambda=0$, then
    $a_i=0$ for all $i\in\mathbb{N}$, implying $v=0$, a contradiction. Thus
    $\lambda\neq0$ and $a_0=0$. But this would also give
    $a_{i+1}=a_i/\lambda=0$ for all $i\in\mathbb{N}$, which implies $v=0$,
    also a contradiction. Thus no such $v$ exists.
  \end{proof}
\end{document}
