\documentclass{article}
\usepackage[left=3cm,right=3cm,top=3cm,bottom=3cm]{geometry}
\usepackage{amsmath,amssymb,amsthm}
%\usepackage{pgfplots,tikz,tikz-cd}
\usepackage[inline]{enumitem}
\usepackage{color}
\setlength{\parindent}{0mm} %So that we do not indent on new paragraphs
\newcommand{\TODO}[1]{\textcolor{red}{TODO: #1}}

\begin{document}
\title{Graduate Algebra II: Homework 9}
\author{Li Ling Ko\\ lko@nd.edu}
\date{\today}
\maketitle

\it \textbf{DF 13.6.3:} Prove that if a field contains the $n$-th roots of
  unity for $n$ odd then it also contains the $2n$-th roots of unity.

  \begin{proof}
    We first show that for $n$ odd,
    \[\mu_{2n} =\mu_n \cup \{-\zeta:\zeta\in\mu_n\}.\]
    Clearly $\supseteq$ holds since $x^n=1$ implies $x^{2n}=1=(-x)^{2n}$.
    For $\subseteq$, observe that if $x^n=1$, then $(-x)^n=-x^n\neq1$.
    Therefore the right side of the equality has exactly $2n$ elements.
    Then since $|\mu_{2n}|=2n$, the equality above holds. \\

    Then since fields are closed under negation,
    $\mathbb{Q}(\mu_n)=\mathbb{Q}(\mu_{2n})$ as required.
  \end{proof}

\it \textbf{DF 13.6.4:} Prove that if $n=p^km$ where $p$ is a prime and $m$
  is relatively prime to $p$ then there are precisely $m$ distinct $n$-th
  roots of unity over a field of characteristic $p$.

  \begin{proof}
    Over a field of characteristic $p$, by Proposition 35,
    \[x^{p^km}-1 =(x^m-1)^{p^k},\]
    so there are at most $m$ distinct $m$-th roots of unity. It suffices to
    show that the roots of $x^m-1$ are distinct. Differentiating with
    respect to $x$ gives $D_x(x^m-1)=mx^{m-1}$, which is non-zero since $p$
    does not divide $m$. Therefore for $m>1$, $x^m-1$ and its derivative
    are coprime, so the roots of $x^m-1$ are distinct by Proposition 33.
  \end{proof}

\it \textbf{DF 13.6.5:} Prove that there are only a finite number of roots
  of unity in any finite extension $K$ of $\mathbb{Q}$.

  \begin{proof}
    If $\zeta_n$ lies in $K$, then the degree $\varphi(n)$ of its cyclotomic
    polynomial $\Phi_n(x)$ cannot be larger than $[K:\mathbb{Q}]$.
    Therefore it suffices to that given any $m\in\mathbb{N}$, there can
    only be a finite number of $m\in\mathbb{N}$ such that $\varphi(m)=n$.
    \\

    To that end, fix $m\in\mathbb{N}$. Let $p_1<\ldots<p_m$ be the
    first $m$ primes. For each prime $p_i$, let $k_i\in\mathbb{N}$ be the
    smallest positive integer such that $p_i^{k_i}-p_i^{k_i-1}>m$. Let $n$
    be any positive integer greater than
    \[N:=(p_1^{k_1}-p_1^{k_1-1}) \cdots (p_m^{k_m}-p_m^{k_m-1}).\]

    Then either $n$ has a prime factor $p$ smaller than $p_m$ such that
    $p^{k}|n$ for some $k>k_r-1$, or $n$ has at least one prime factor $q$
    greater than $p_m$. In the former case, $p^k-p^{k-1}$ divides
    $\varphi(n)$, so $\varphi$ will be greater than $m$. In the latter
    case, the prime factor $q$ that is greater than $p_m$ must also be
    greater than $m$, since $p_m>m$. But $q^k-q^{k-1}$ divides $\varphi(n)$
    for some $k\geq1$, so $\varphi$ is also greater than $m$. Thus all
    integers greater than $N$ will have Euler-totient value greater than
    $m$.
  \end{proof}

\it \textbf{DF 13.6.6:} Prove that for $n$ odd, $\Phi_{2n}(x)=\Phi_n(-x)$.
  \begin{proof}
    In Exercise 13.6.3, we proved that when $n$ is odd,
    \[\mu_{2n} =\mu_n \cup \{-\zeta:\zeta\in\mu_n\}.\]

    Therefore the $2n$-th roots of unity either satisfy $\Phi_n(x)$ or
    $\Phi_n(-x)$. Note that $\Phi_n(-x)$ is a monic polynomial because
    $\text{deg}(\Phi_n(x)) =\varphi(n)$ which is even because $n$ is odd.
    Now since $\Phi_n(x)$ is irreducible over $\mathbb{Q}$, $\Phi_n(-x)$
    must also be irreducible. Therefore $\Phi_{2n}(x)$ is either
    $\Phi_n(x)$ or $\Phi_n(-x)$. Now the roots of $\Phi_n(x)$ cannot
    contain the primitive $2n$-th root of unity since it contains exactly
    the $n$-th roots of unity, therefore the primitive $2n$-th root of
    unity must satisfy the other polynomial $\Phi_n(-x)$. Therefore
    $\Phi_{2n}(x)=\Phi_n(x)$.
  \end{proof}

\it \textbf{DF 13.6.8:} Let $l$ be a prime and let $\Phi_l(x)
  =\frac{x^l-1}{x-1} =x^{l-1}+x^{l-2}+\ldots+x+1 \in\mathbb{Z}[x]$ be the
  $l$-th cyclotomic polynomial, which is irreducible over $\mathbb{Z}$ by
  Theorem 41. This exercise determines the factorization of $\Phi_l(x)$
  modulo $p$ for any prime $p$. Let $\zeta$ denote any fixed primitive $l$-th
  root of unity.

  \begin{enumerate}[label={(\alph*)}]
    \item Show that if $p=l$ then $\Phi_l(x)=(x-1)^{l-1}
      \in\mathbb{F}_l[x]$.

      \begin{proof}
        Over $\mathbb{Q}$ and therefore also over $\mathbb{F}_l$, we have
        \[(x-1)\Phi_l(x) =x^l-1.\]
        Now over $\mathbb{F}_l$, the right hand side is $x^l-1=(x-1)^l$.
        Then dividing by the $x-1$ term on both sides give
        \[\Phi_l(x)=(x-1)^{l-1} \in\mathbb{F}_l.\]
      \end{proof}

    \item Suppose $p\neq l$ and let $f$ denote the order of $p\mod l$, i.e.
      $f$ is the smallest power of $p$ with $p^f\equiv1\mod l$. Use the
      fact that $\mathbb{F}_{p^n}^\times$ is a cyclic group to show that
      $n=f$ is the smallest power $p^n$ of $p$ with
      $\zeta\in\mathbb{F}_{p^n}$. Conclude that the minimal polynomial of
      $\zeta$ over $\mathbb{F}_p$ has degree $f$.

      \begin{proof}
        Let $n\in\mathbb{N}$ a positive integer such that
        $\zeta\in\mathbb{F}_{p^n}$. Since $\mathbb{F}_{p^n}^\times$ is a
        cyclic group, it must be generated by some element $a$ with order
        (as a group under multiplication) $p^n-1$. Then $\zeta=a^k$ for some
        $k<p^n$. Now $\zeta^l=1$, so $a^{kl}=1$. But $a$ has order $p^n-1$,
        so $p^n-1$ must divide $kl$, which implies $p^n\equiv1\mod l$. \\

        Conversely, if $p^n\equiv1\mod l$, then $p^n-1=kl$ for some
        $k\in\mathbb{N}$. Then in the group $\mathbb{F}_{p^n}^\times$ with
        generator $a$, element $a^k$ will satisfy $x^l=x^{p^n-1}=1$.
        Therefore $\mathbb{F}_{p^n}^\times$ will contain $\zeta$. \\

        Since $\mathbb{F}_{p^n}$ contains $\zeta$ if and only if
        $p^n\equiv1\mod l$, the given $n=f$ is the smallest power $p^n$ of
        $p$ with $\zeta\in\mathbb{F}_{p^n}$. \\

        Therefore, $\mathbb{F}_{p^f}=\mathbb{F}_p(\zeta)$, so the degree of
        the minimal polynomial of $\zeta$ over $\mathbb{F}_p$ is
        \[[\mathbb{F}_p(\zeta):\mathbb{F}_p] =[\mathbb{F}_{p^f}:\mathbb{F}_p]
        =f.\]
      \end{proof}

    \item Show that $\mathbb{F}_p(\zeta)=\mathbb{F}_p(\zeta^a)$ for any
      integer $a$ not divisible by $l$. Conclude using Part (b) that, in
      $\mathbb{F}_p[x]$, $\Phi_l(x)$ is the product of $\frac{l-1}{f}$
      distinct irreducible polynomials of degree $f$.

      \begin{proof}
        Clearly $\mathbb{F}_p(\zeta) \supseteq\mathbb{F}_p(\zeta^a)$. For
        the reverse inclusion, since $l$ is prime, if $a$ is not divisible
        by $l$, then $(a,l)=1$, and $\zeta^a$ is also a primitive $l$-th
        root of unity. In particular, $\zeta=(\zeta^a)^k$ for some integer
        $k$, therefore $\mathbb{F}_p(\zeta)
        \subseteq\mathbb{F}_p(\zeta^a)$. \\

        Because $l$ is prime, there are exactly $(l-1)$ $l$-th roots of
        unity, given by $\zeta,\zeta^2,\ldots,\zeta^{l-1}$. These roots are
        the $l-1$ distinct roots of $\Phi_l(x)$. Since the irreducible
        polynomial of each of these roots has degree $f$ given in part (b),
        $\Phi_l(x)$ must be the product of $\frac{l-1}{f}$ distinct
        irreducible polynomials of degree $f$. Note that the polynomials
        must be distinct otherwise the roots would of $\Phi_l(x)$ would
        repeat.
      \end{proof}

    \item In particular, prove that, viewed in $\mathbb{F}_p[x]$,
      $\Phi_7(x) =x^6+x^5+\ldots+x+1$ is $(x-1)^6$ for $p=7$, a product of
      distinct linear factors for $p\equiv1\mod 7$, a product of 3
      irreducible quadratics for $p=6\mod7$, a product of 2 irreducible
      cubics for $p=2,4\mod7$, and is irreducible for $p=3,5\mod7$.

      \begin{proof}
        Given that $l=7$, we first find the smallest $f\in\mathbb{N}$ such
        that $p^f\equiv1\mod l$ for the given $p$'s:
        \begin{align*}
          1^1\equiv 1\mod7 \\
          2^3\equiv 1\mod7 \\
          3^6\equiv 1\mod7 \\
          4^3\equiv 1\mod7 \\
          5^6\equiv 1\mod7 \\
          6^2\equiv 1\mod7 \\
        \end{align*}

        Therefore from part (c), we get the assertions given in the
        question.
      \end{proof}
  \end{enumerate}

\it \textbf{DF 13.6.9:} Suppose $A$ is an $n\times n$ matrix over
  $\mathbb{C}$ for which $A^k=I$ for some integer $k\geq1$. Show that $A$
  can be diagonalized. Show that the matrix $A=\begin{pmatrix} 1&\alpha\\
  0&1\\ \end{pmatrix}$ where $\alpha$ is an element of a field of
  characteristic $p$ satisfies $A^p=I$ and cannot be diagonalized if
  $\alpha\neq0$.

  \begin{proof}
    Let $A$ be an $n\times n$ matrix over $\mathbb{C}$ for which $A^k=I$
    for some integer $k\geq1$. Since this is a matrix over $\mathbb{C}$,
    all the roots of its characteristic polynomial will be contained in
    $\mathbb{C}$. Also, matrix $A$ satisfies the polynomial
    $x^k-1=0$, which has distinct roots by Proposition 33, therefore the
    minimal polynomial of $A$ must have distinct roots. Then from
    Corollary 12.3.25, $A$ must be diagonalizable. \\

    Let $A=\begin{pmatrix} 1&\alpha\\ 0&1\\ \end{pmatrix}$ over
    $\mathbb{F}_p$ and where $\alpha$ is an element of a field of
    characteristic $p$. By induction on $n$, we have
    \[A^n =\begin{pmatrix} 1&n\alpha\\ 0&1\\ \end{pmatrix}.\]

    Therefore $A^p=I$. Since $\alpha\neq0$, the matrix $B=A/\alpha$ is
    well-defined, and is a Jordan block for the eigenvalue $1/\alpha$.
    Then by uniqueness of the Jordan canonical form (Theorem 12.3.23),
    $A$ is not diagonalizable.
  \end{proof}

\it \textbf{DF 13.6.10:} Let $\varphi$ denote the Frobenius map $x\mapsto
  x^p$ on the finite field $\mathbb{F}_{p^n}$. Prove that $\varphi$ gives
  an isomorphism of $\mathbb{F}_{p^n}$ to itself. A Prove that $\varphi^n$
  is the identity map and that no lower power of $\varphi$ is the identity.

  \begin{proof}
    Clearly $\varphi$ preserves addition and multiplication from
    Proposition 35, thus it is a ring homomorphism.
    Recall that $\mathbb{F}_{p^n}^\times \cong\mathbb{Z}_{p^n-1}^+$.
    Then $\varphi$ must be injective since homomorphism of fields are
    injective. Then since $\mathbb{F}_{p^n}$ is finite, $\varphi$ must also
    be surjective, therefore $\varphi$ is an automorphism. \\

    $\varphi^n$ maps $x$ to $x^{p^n}$. Then since $\mathbb{F}_{p^n}^\times
    \cong\mathbb{Z}_{p^n-1}^+$, all non-zero elements in $\mathbb{F}_{p^n}$
    have multiplicative order $p^n-1$, therefore $\varphi$ is the identity
    map. \\

    Any lower power $m$ of $\varphi$ will map $x$ to $x^{p^n}$. Let
    $a\in\mathbb{F}_{p^n}^\times$ be a generator. Then
    $\varphi^m(a)=a^{p^m}$, which will equal to $a$ for the first time when
    $p^m-1$ equals $p^n-1$ for the first time, since $a$ has multiplicative
    order $p^n-1$. Therefore $n$ is the lowest such that $\varphi^n$ gives
    the identity. 
  \end{proof}
\end{document}
