\documentclass{article}
\usepackage[left=3cm,right=3cm,top=3cm,bottom=3cm]{geometry}
\usepackage{amsmath,amssymb,amsthm,tikz,mathtools}
\usepackage{color}
\usepackage[inline]{enumitem}
\usetikzlibrary{patterns}
\setlength{\parindent}{0mm}
\newcommand{\TODO}[1]{\textcolor{red}{TODO: #1}}

\begin{document}
\title{Basic Logic II: Problem Set 1}
\author{Li Ling Ko\\ lko@nd.edu}
\date{\today}
\maketitle

\begin{enumerate}[label={\bf Q\arabic*:}]
  \item \it Show the relation $R(x,y,p):=\{(x,y,p)|xy\neq p\}$ is primitive
    recursive.

    \begin{proof}
      Note that
      \[\chi_R(\bar{x})=\chi_{\neq}(\text{mult}(\bar{x}_1,
        \bar{x}_2), \bar{x}_3),\]
      where $\text{mult}(x_1,x_2)$ denotes the multiplication function,
      $\bar{x}_i$ denotes the $i$th projection function, and
      $\chi_{\neq}(x_1,x_2)$ denotes the graph of the inequality relation.
      Thus by closure under projection and composition, it suffices to show
      that the constituent functions $\text{mult}(x_1,x_2)$ and
      $\chi_{\neq}(x_1,x_2)$ are primitive recursive. \\

      The multiplication function can be derived as:
      \[\begin{array}{lcl}
        \text{mult}(0,x_2) &= &0, \\
        \text{mult}(x_1+1,x_2) &= &\text{mult}(x_1,x_2)+x_2 \\
        &= &\text{add}(\text{mult}(x_1,x_2),x_2) \\
        &= &h(x_1,\text{mult}(x_1,x_2),x_2), \\
      \end{array}\]
      where $h(\bar{x}):=\text{add}(\bar{x}_2,\bar{x}_3)$. Now the zero
      function is primitive recursive by definition. Also, $h(\text{x})$ is
      primitive recursive by closure under projection and composition, and
      also by the fact that the addition function $\text{add}(x_1,x_2)$ is
      primitive recursive (example on page 2 of notes). Therefore,
      $\text{mult}(x_1,x_2)$ is primitive recursive by closure under
      primitive recursion. \\

      It remains to show that the inequality relation is primitive
      recursive.
    \end{proof}

  \item \it What is the code of the primitive recursive used in your third
    step of the above derivation?

    \begin{proof}
    \end{proof}

  \item \it Find a ``computable'' function $f$ that dominates all primitive
    recursive functions.

    \begin{proof}
    \end{proof}
\end{enumerate}
\end{document}
