\documentclass{article}
\usepackage[left=3cm,right=3cm,top=3cm,bottom=3cm]{geometry}
\usepackage{amsmath,amssymb,amsthm,tikz,mathtools}
\usepackage{color}
\usepackage[inline]{enumitem}
\usetikzlibrary{patterns}
\setlength{\parindent}{0mm}
\newcommand{\TODO}[1]{\textcolor{red}{TODO: #1}}

\begin{document}
\title{Basic Logic II: Problem Set 3}
\author{Li Ling Ko\\ lko@nd.edu}
\date{\today}
\maketitle

\begin{enumerate}[label={\bf Q\arabic*:}]
  \item \it Working in $\mathbb{N}$ show that for every $k\in\mathbb{N}$ and
    for every subset $F\subseteq[0,k]$, there are $m,n\in\mathbb{N}$ such
    that for every $i\leq k$, $i\in F \Leftrightarrow m(i+1)+1|n$.

    \begin{proof}
      Choose $m=k!$. We first show that $m+1,\ldots,m(k+1)+1$ are pairwise
      coprime: Assume that $mx+1$ and $my+1$ are not coprime for some
      distinct $x,y\in\{1,\ldots,k+1\}$. We can assume without loss of
      generality that $x>y$. Let $p$ be the smallest prime that divides
      $mx+1$ and $my+1$. Then $p$ must also divide $(mx+1)-(my+1)=m(x-y)$.
      But $x-y\in\{1,\ldots,k\}$ and $m=k!$, so $p$ must divide one of
      $\{2,\ldots,k\}$, and therefore divide $mx$ and $my$. Thus $p$,
      cannot divide $mx+1$ and $my+1$, a contradiction. \\

      So the integers $m+1,\ldots,m(k+1)+1$ are pairwise coprime. Consider
      the system of $k+1$ equations in $n$:
      \[\begin{array}{llll}
        n &\equiv &\delta(0\not\in F) &\mod{m+1}\\
        n &\equiv &\delta(1\not\in F) &\mod{2m+1}\\
        \ldots &&&\\
        n &\equiv &\delta(k\not\in F) &\mod{(k+1)m+1},\\
      \end{array}\]
      where
      \[\delta(x)
      :=\begin{cases}1,\text{if}\;x,\\0,\text{otherwise}\end{cases}.\]

      By the Chinese remainder theorem, there is a solution for $n$. This
      $n$ will satisfy the requirement we are looking for.
    \end{proof}

  \item \it Let $\varphi$ be $\sum_1$ in $\mathbb{N}$. Consider in
    $\mathbb{N}$ the definable set $W=\{a|\varphi(a)\}$. Is $W$ computably
    enumerable? Why or why not?

    \begin{proof}
      Yes $W$ is computably enumerable. If $W$ is finite then it is
      trivially computably enumerable. So assume $W$ is infinite. Now since
      $\varphi$ is $\sum_1$, $\varphi(x)$ is equivalent to $\exists
      y\;\phi(x,y)$ for some $\Delta_0$-formula $\phi$. Fix any
      computable permutation
      $p:\mathbb{N}\rightarrow\mathbb{N}\times\mathbb{N}$, and let $\pi_i$
      denote the $i$th projection function. Consider the function
      $f(n):\mathbb{N}\rightarrow\mathbb{N}$ defined recursively as
      follows: \\

      At input $0$, we find the first $s_0\in\mathbb{N}$ such that
      $\phi(\pi_1(p(s_0)),\pi_2(p(s_0)))$ holds, and output
      $\pi_1(p(s_0))$. Then, at input $n+1$, we find the next
      $s_{n+1}>s_{n}$ such that $\phi(\pi_1(p(s_{n+1})),\pi_2(p(s_{n+1})))$
      holds, and output $\pi_1(p(s_{n+1}))$. Note that the $s_{n}$'s exists
      because $W$ is infinite. Also, given $m,n\in\mathbb{N}$, checking if
      $\phi(m,n)$ holds is al algorithmic procedure that always terminates
      since $\phi$ is $\Delta_0$.  So $f$ is defined by an
      algorithmic procedure and produces an output for every input,
      therefore it is total recursive by the Church-Turing thesis. Also,
      since $p$ is surjective, $f$ will enumerate all elements of $W$. 
    \end{proof}

  \item \it Assume $\varphi(x,y)$ is $\Delta_1$ in $\mathbb{N}$ and
    in $\mathbb{N}$, $\varphi$ is total (so for all $x$ there is a unique
    $y$ such that $\varphi(x,y)$). As a function is $\varphi$ computable?
    Why or why not?

    \begin{proof}
      Given a formula $\phi(\bar{x})$, let $\phi(\mathbb{N})$ denote
      $\{\bar{n}\in\mathbb{N}^n|\phi(\bar{n})\}$. We have shown in
      question 2 that $\varphi(\mathbb{N})$ is computably enumerable since
      $\varphi$ is $\sum_1$. Hence for a given $x_0\in\mathbb{N}$, we
      enumerate the $\varphi(\mathbb{N})$'s to find the first tuple where
      the first coordinate is $x_0$, and output the second coordinate.
      This procedure is computable by the Church-Turing thesis, and
      describes $\varphi$ as a function. \\

      Comment: I don't think $\varphi$ needs to be $\prod_1$ for $\varphi$
      to be computable as a function.
    \end{proof}

  \item \it Show that for every computably enumerable theory $T$ there is
    a computable theory $T^*$ such that
    \[\{\varphi:T\vdash\varphi\} = \{\varphi:T^*\vdash\varphi\}.\]

    \begin{proof}
      Let $f:\mathbb{N}\rightarrow\mathcal{L}\text{-formulas}$ be an
      injective and total recursive function that enumerates $T$, and
      consider
      \[T^*:=\{f(0)\wedge\ldots\wedge f(n): n\in\omega\}.\]

      Then clearly $\{\varphi:T\vdash\varphi\}
      =\{\varphi:T^*\vdash\varphi\}$. To check if a given formula $\phi$
      lies in $T^*$, we first decompose $\phi$ into its finite conjunction
      of formulas $\phi=\phi_0\wedge\ldots\wedge\phi_n$; this decomposition
      is unique if the coding is done properly. Next, we check if each
      $\phi_i$ is the same as $f(i)$, and decide $\phi\in T^*$ if and only
      if $\phi_i=f(i)$ for all $i\in\{0,\ldots,n\}$. Since the described
      procedure is algorithmic, by the Church-Turing thesis, $T^*$ is
      computable.
    \end{proof}

  \item \it A theory is complete if and only if for all sentences
    $\varphi$, $T\vdash\varphi$ or $T\vdash\neg\varphi$. Show $Q$ is is not
    complete. A similar proof should show that $Q$ plus induction is not
    complete.

    \begin{proof}
      Assume $Q$ is complete. We show that because the axioms of $Q$ is
      recursively enumerable, the theory of $Q$ would be decidable,
      contradicting Church's Theorem (Corollary 68). Since the axioms of
      $Q$ is enumerable, we can enumerate all possible proofs from the
      axioms, and thus also enumerate all theories of $Q$. So to decide if
      a given any sentence $\varphi$ is a theory of $Q$, we enumerate all
      the $Q$ theories until we reach either $\varphi$ or $\neg\varphi$; we
      will always reach one or the other since $Q$ is complete by
      assumption. Thus the theory of $Q$ is decidable, a contradiction.
    \end{proof}

  \item \it A theory is decidable if and only if
    $\{\varphi:T\vdash\varphi\}$ is computable. Let $T$ be a computable
    theory with finitely many complete extensions. Show that $T$ is
    decidable.

    \begin{proof}
      Note that we can assume that $T$ is consistent, otherwise $T$ is
      trivially decidable. \\

      Let $T_1,\ldots,T_n$ be the distinct complete extensions of $T$.
      We first show that for each $i\in\{1,\ldots,n\}$, the theory of $T_i$
      is the same as the theory of $T\cup S_i$ for some finite set of
      sentences $S_i\subset T_i$: For each $1\neq j\in\{1,\ldots,n\}$,
      since $T_i$ is different from $T_j$, it must contain some sentence
      $\varphi_{i,j}$ that is not contained in $T_j$, i.e.
      $\neg\varphi_{i,j}\in T_j$. Choose
      \[S_i :=\{\varphi_{i,1},\ldots,\varphi_{i,i-1},
      \varphi_{i,i+1},\ldots,\varphi_{i,n}\}.\]

      To show that $T\cup S_i$ is complete, assume it is not. Then there
      must be a sentence $\varphi$ such that both $T\cup
      S_i\cup\{\varphi\}$ and $T\cup S_i\cup\{\neg\varphi\}$ are
      consistent. Let $T_i'$ and $T_i''$ be respective complete extensions
      of those theorems. Then from choice of $S_i$, neither $T_i'$ nor
      $T_i''$ can be $T_j$ for any $j\neq i$, since $S_i\in T_i'$ and
      $S_i\in T_i''$. Thus $T_i'$ and $T_i''$ are both equal to $T_i$,
      which means $T_i$ contains both $\varphi$ and $\neg\varphi$,
      contradicting consistency of $T_i$. \\

      To decide if a given any sentence $\varphi$ is a consequence of $T$,
      we use the fact $\varphi$ is a consequence of $T$ if and only if it
      is a consequence of all possible extensions of $T$, by Godel's
      completeness theorem. In other words, $T\vdash\varphi$ if and only if
      $T_i\vdash\varphi$ for all $i\in\{1,\ldots,n\}$. Thus it suffices to
      show that each $T_i$ is decidable. Now since the theory of $T_i$ is
      the same as the theory of $T\cup S_i$, it suffices to show that
      $T\cup S_i$ is decidable. To this end, note that $T$ is enumerable
      and $S_i$ is finite, so $T\cup S_i$ is also enumerable. Also, $T\cup
      S_i$ is complete since it is the same theory as $T_i$, thus from the
      claim made in question 5 (where we showed that recursively enumerable
      and complete theories are decidable), $T_i$ is decidable, which
      completes our proof.
    \end{proof}
\end{enumerate}
\end{document}
