\documentclass{article}
\usepackage[left=3cm,right=3cm,top=3cm,bottom=3cm]{geometry}
\usepackage{amsmath,amssymb,amsthm,pgfplots,tikz}
\usetikzlibrary{patterns}
\usepackage{color}
\setlength{\parindent}{2mm}
\newcommand{\TODO}[1]{\textcolor{red}{TODO: #1}}

\begin{document}
\title{Geometry and Topology: Homework 2}
\author{Li Ling Ko\\ lko@nd.edu}
\date{\today}
\maketitle

\begin{enumerate}
  \item Let $X$ be a CW complex.
    \begin{enumerate}
      \item Prove that if $X$ has finitely may cells, then $X$ is compact.
        \begin{proof}
          We prove by induction on $n$, the number of cells of $X$. For the
          base case where $n=1$, $X$ is a single point, which is trivially
          compact. For the inductive step, assume the statement is true for
          $n$ cells, and consider the case when $X$ is composed of $(n+1)$
          cells. $X$ is obtained from gluing an $n$-disc $D^n$ to $X^{(n)}$
          at the boundary of the disk $\delta D^n$. $X^{(n)}$ is compact by
          inductive assumption. The $n$-disc is also compact since it is
          closed and bounded in $\mathbb{R}^n$. Since $X$ is a compact
          space attached to another compact space, it must also be compact
          - any open cover of $X$ must cover each of its component
          spaces, giving a finite cover from each component space by the
          compactness of those spaces.
        \end{proof}

      \item Let $C\subset X$ be a compact subset. Prove that $C$ only
        intersects finitely many cells of $X$.
        \begin{proof}
          Assume by contradiction that a compact subset $C$ intersects
          infinite number of cells of $X$. Let $\{e^i\}_{i\in\mathbb{N}}$
          be a countably infinite series of distinct cells that intersect
          with $C$. From each cell $e^i$, we pick a point $x_i\in X$ from
          their area of intersection, to obtain a series $P=\{x_0,\ldots\}$
          of points in $X$. \\

          We first prove by induction that $P$ is closed in $X$. By
          definition, this is equivalent to proving that $P\cap X^{(n)}$ is
          closed in $X^{(n)}$ for each $n\in\mathbb{N}$. We prove this
          claim by induction on $n$: The base case is trivially true since
          it consists $P\cap X^{(0)}$ consists of only one point $x_0$.
          Assume the claim is true for $n$. Then $\varphi_{n+1}^{-1}(P)$ is
          closed in the disk boundary $\delta D^{n+1}$, so after adding one
          more point $x_{n+1}$ into $D^{n+1}$, $\varphi_{n+1}^{-1}(P)$ will
          remain closed in $D^{n+1}$, implying that $P\cap X^{n+1}$ is
          closed in $X^{n+1}$. \\

          By similar argument, any subset of $P$ must also be closed in
          $X$, which means that $P$ as a subset of $X$ has the discrete
          topology. Yet, $P$ is infinite, so it cannot be compact, which is
          a contradiction because $P$ is a closed subset of a compact space
          $C$, and should be compact.
        \end{proof}
    \end{enumerate}

  \item Construct CW complex structures on the following spaces.
    \begin{enumerate}
      \item An $n$-dimensional torus.
        \begin{proof}
          An $n$-torus is formed by first creating an $n$-dimensional cube,
          then gluing opposing faces of the cube together. We observe the
          pattern of number cells needed as $n$ increases from 0 to 2. The
          0-torus is a single 0-cell. The 1-torus is a 0-cell, glued with a
          1-cell. The 2-torus is a 0-cell, glued with two 1-cells, then
          with a 2-cell on the two one-cells. So the $n$-cell is obtained
          by starting with a 0-cell, then gluing $n$ 1-cells to the 0-cell,
          where the end points of each 1-cell is glued to the 0-cell,
          forming a flower with $n$ petals. To complete the $n$-cell, we
          build a $(n-1)$-torus over each $\binom{n}{n-1}$ choice of
          $(n-1)$ petals, reusing any $(n-i)$ cells that have been built
          earlier. \\

          This system of construction can be summarized as follows
          \begin{enumerate}
            \item Start with a single 0-cell.
            \item Glue $n$ 1-cells to the 0-cell, sticking the end points
              of each 1-cell to the 0-cell, forming a flower with $n$
              petals.
            \item Repeat for $i=2$ to $n$: Between every $i$
              1-cells, glue an $i$-cell to the $i$ $(i-1)$-cells
              constructed in the earlier stage, such that opposing faces of
              the $i$-cell are glued to the same $(i-1)$-cell, and each
              $(i-1)$-cell has exactly two $i$-cell faces glued onto them.
          \end{enumerate}

          From construction, we summarize that an $n$-torus is built from
          $\binom{n}{i}$ $i$-cells, for $0\leq i\leq n$. \\
        \end{proof}

      \item Letting $\{p_1,\ldots,p_n\}$ be $n$ distinct points on $S^2$,
        the quotient space of $S^2$ that identifies all the $p_i$ to a
        single point.

        \begin{proof}
          After visualizing, we realize that the quotient space of
          $S^2$ is constructed using one 0-cell, $(n-1)$ 1-cells, and one
          2-cell, in the following manner:
          \begin{enumerate}
            \item Start with a 0-cell, which is a single point.
            \item Glue the $(n-1)$ 1-cells to the 0-cell, sticking the end
              points of each 1-cell to the 0-cell, forming a flower with
              $(n-1)$ petals labelled $a_1$ to $a_{n-1}$, as shown in the
              left figure below.
            \item Partition the boundary of the 2-cell into $2(n-1)$ parts,
              then label and direct the parts as shown in the right figure
              below. Then, glue the 2-cell onto the petals of the flower
              according to the labels, sending each of the marked dots to
              the 0-cell.
          \end{enumerate}

          \begin{center}
            \begin{tikzpicture}
              %\draw[help lines](0,0) grid (12,6);
              \foreach\Point in {
                (3,3),
                (9.5,1), (9.5,5),
                (8.5,2), (8.5,3), (8.5,4),
                (10.5,2), (10.5,3), (10.5,4)
                }{\node at \Point {\textbullet};}
              \path[draw,dotted,line width=1.5pt] (8.5,1)--(10.5,1);
              \path[draw,line width=1.5pt] (8.5,5)--(10.5,5);
              \draw[line width=1.5pt,dotted] (10.5,1) arc (-90:0:0.5);
              \draw[line width=1.5pt,dotted] (10.5,2) arc (-90:0:0.5);
              \draw[line width=1.5pt] (10.5,3) arc (-90:0:0.5);
              \draw[line width=1.5pt] (10.5,4) arc (-90:0:0.5);

              \draw[line width=1.5pt,dotted,<-] (11,1.5) arc (0:90:0.5);
              \draw[line width=1.5pt,dotted,<-] (11,2.5) arc (0:90:0.5);
              \draw[line width=1.5pt,<-] (11,3.5) arc (0:90:0.5);
              \draw[line width=1.5pt,<-] (11,4.5) arc (0:90:0.5);

              \draw[line width=1.5pt,dotted,->] ( 8.5,2) arc (90:180:0.5);
              \draw[line width=1.5pt,dotted,->] ( 8.5,3) arc (90:180:0.5);
              \draw[line width=1.5pt,->] ( 8.5,4) arc (90:180:0.5);
              \draw[line width=1.5pt,->] ( 8.5,5) arc (90:180:0.5);

              \draw[line width=1.5pt,dotted] (8,1.5) arc (180:270:0.5);
              \draw[line width=1.5pt,dotted] (8,2.5) arc (180:270:0.5);
              \draw[line width=1.5pt] (8,3.5) arc (180:270:0.5);
              \draw[line width=1.5pt] (8,4.5) arc (180:270:0.5);

              \node[] at (11.5,1.5) {$a_{n-1}$};
              \node[] at (11.5,3.5) {$a_2$};
              \node[] at (11.5,4.5) {$a_1$};

              \node[] at (7.5,1.5) {$a_{n-1}$};
              \node[] at (7.5,3.5) {$a_2$};
              \node[] at (7.5,4.5) {$a_1$};
              \node[] at (9.5,0.5) {2-cell};

              \path[draw,line width=1.5pt,dotted] (2.5,1)--(3,3)--(3.5,1);
              \path[draw,line width=1.5pt] (2.5,5)--(3,3)--(3.5,5);
              \path[draw,line width=1.5pt,dotted] (5,3.5)--(3,3)--(5,2.5);
              \path[draw,line width=1.5pt] (1,3.5)--(3,3)--(1,2.5);

              \draw[line width=1.5pt,->] (2.5,5) arc (180:90:0.5);
              \draw[line width=1.5pt] (3,5.5) arc (90:0:0.5);
              \draw[line width=1.5pt,-<,dotted] (2.5,1) arc (-180:-90:0.5);
              \draw[line width=1.5pt,dotted] (3,0.5) arc (-90:0:0.5);
              \draw[line width=1.5pt,->,dotted] (5,3.5) arc (90:0:0.5);
              \draw[line width=1.5pt,dotted] (5.5,3) arc (0:-90:0.5);
              \draw[line width=1.5pt,->] (1,2.5) arc (-90:-180:0.5);
              \draw[line width=1.5pt] (0.5,3) arc (180:90:0.5);

              \node[] at (0,3) {$a_{1}$};
              \node[] at (3,6) {$a_{2}$};
              \node[] at (3,0) {$a_{n-1}$};
            \end{tikzpicture}
          \end{center}
        \end{proof}
    \end{enumerate}

  \item Carefully prove that the following are covering spaces.
    \begin{enumerate}
      \item
        \begin{enumerate}
          \item The map $\pi:\mathbb{C}\rightarrow\mathbb{C}^*$ defined by
            $\pi(z)=e^z$. \label{qn:exp}

            \begin{proof}
              First, we observe that the function is holomorphic, and hence
              is open and continuous by the open mapping theorem. Also,
              given any $y\in\mathbb{C}^*$, the inverse $z=\ln(y)$ always
              exists, and $\pi(z_1)=\pi(z_2)\leftrightarrow z_1-z_2\in
              2\pi i\mathbb{Z}$. Hence, the map when restricted to any open
              subset of the horizontal strip $\mathbb{R}\times[z-\pi
              i,z+\pi i)\subset\mathbb{C}^*$, where $z\in\mathbb{C}$, is
              a homeomorphism. We use these properties to show that given any
              $y\in\mathbb{C}^*$, there is an open set $\mathcal{O}$
              containing $y$ such that its pre-image is composed of
              disjoint open sets each of which is homeomorphic to
              $\mathcal{O}$. \\

              Given $y\in\mathbb{C}^*$, choose any $z\in\mathbb{C}$ in
              its pre-image, and let $\mathcal{U}\in\mathbb{C}$ be an open
              ball $B_r(z)$ of radius $0<r<0.5$ around $z$. Let $\mathcal{O}$
              be the the image of $\mathcal{U}$ under $\pi$; this image
              will contain $y$ and is open because $\pi$ is an open
              function. Consider the restriction of $\pi$ to the open ball
              $\mathcal{U}$. Since the diameter of $\mathcal{U}$ is less
              than 1, the map is injective. Hence, by the holomorphic
              property of $\pi$, $\mathcal{U}$ is homeomorphic to
              $\mathcal{O}$. Now the pre-image of $\mathcal{O}$ is the
              union of $2\pi ki$ translations of $\mathcal{U}$ for
              $k\in\mathbb{Z}$, or more precisely
              
              \begin{equation*}
                \pi^{-1}(\mathcal{O}) = \bigcup\limits_{k\in\mathbb{Z}}
                (\mathcal{U}+2k\pi i),
              \end{equation*}

              where each $\mathcal{U}+2\pi ki=\{x+2k\pi
              i:x\in\mathcal{U}\}$ is open, homeomorphic to
              $\mathcal{O}$, and pairwise disjoint, which completes the
              proof.
            \end{proof}

          \item For $n\in\mathbb{Z}\setminus\{0\}$, the map
            $\pi:\mathbb{C}^*\rightarrow\mathbb{C}^*$ defined by
            $\pi(z)=z^n$.

            \begin{proof}
              Similar to the proof in question~\ref{qn:exp}, observe that
              the map is holomorphic and is therefore open and continuous
              by the open mapping theorem. Also, given any
              $y\in\mathbb{C}^*$, the inverse $z=y^{1/n}$ always exists,
              and the pre-image of $y$ are the $n$ points
              $\pi^{-1}(y)=\{y^{1/n}\exp(2k\pi i/n):0\leq k<n\}$. Hence,
              the map when restricted to any open subset of sectors
              $\{z\in\mathbb{C}^*:\theta-\pi i/n<\arg(z)<\theta+\pi i/n\}$,
              where $\theta\in[0,2\pi]$, is a homeomorphism. We use
              these properties to show that given any $y\in\mathbb{C}^*$,
              there is an open set $\mathcal{O}$ containing $y$ such that
              its pre-image is composed of disjoint open sets each of which
              is homeomorphic to $\mathcal{O}$. \\

              Given $y\in\mathbb{C}^*$, choose any $z\in\mathbb{C}$ in
              its pre-image, and let $\mathcal{U}\in\mathbb{C}$ be an open
              ball $B_\epsilon(z)$ around $z$ and contained in the sector
              $\{w\in\mathbb{C}^*:\arg(z)-\pi i/n<\arg(w)<\arg(z)+\pi
              i/n\}$. Such an open ball exists since the sector is open.
              Let $\mathcal{O}$ be the the image of $\mathcal{U}$ under
              $\pi$; this image will contain $y$ and is open because $\pi$
              is an open function. Consider the restriction of $\pi$ to the
              open ball $\mathcal{U}$. Since the difference in $\arg$
              between any two points in $\mathcal{U}$ is less than $2k\pi
              i/n$, the map is injective. Hence, by the holomorphic
              property of $\pi$, $\mathcal{U}$ is homeomorphic to
              $\mathcal{O}$. Now the pre-image of $\mathcal{O}$ is the
              union of $n$ rotations of $\mathcal{U}$ by $2k\pi i/n$
              degrees for $0\leq k<n$, or more precisely

              \begin{equation*}
                \pi^{-1}(\mathcal{O}) = \bigcup\limits_{k=0}^{n-1}
                \mathcal{U}\exp(2k\pi i/n),
              \end{equation*}

              where each $\mathcal{U}\exp(2k\pi i/n)=\{x\exp(2k\pi
              i/n):x\in\mathcal{U}\}$ is open, homeomorphic to
              $\mathcal{O}$, and pairwise disjoint, which completes
              the proof.
            \end{proof}
        \end{enumerate}

      \item Prove that the map $\pi:\mathbb{C}\rightarrow\mathbb{C}$
        defined by $\pi(z)=z^2$ is not a covering space.

        \begin{proof}
          If $\pi$ is a covering space, then given any $z\in\mathbb{C}$ and
          open set $\mathcal{O}$ containing $z$, the cardinality of the
          pre-image of each element in $\mathcal{O}$ should be the same.
          However at $z=0$ and given any open ball $B_\epsilon(0)$ around
          $z$, we have $|\pi^{-1}(0)|=|\{0\}|=1$, but
          $|\pi^{-1}(\epsilon/2)|=|\{-\sqrt{\epsilon/2},\sqrt{\epsilon/2}\}|=2$,
          so $\pi$ cannot be a covering space.
        \end{proof}
    \end{enumerate}

  \item Let $\pi:\tilde{X}\rightarrow X$ be a covering space such that
    $\pi^{-1}(p)$ is finite and nonempty for all $p\in X$. Prove that $X$ is
    compact Hausdorff if and only if $\tilde{X}$ is compact Hausdorff.

    \begin{proof}
      We first prove the forward implication. Assume that $X$ is compact
      Hausdorff. We show that $\tilde{X}$ is Hausdorff. Let
      $\tilde{x_1},\tilde{x_2}$ be two points in $\tilde{X}$. There are two
      cases to consider. In the first case, $\tilde{x_1}$ and $\tilde{x_2}$
      are mapped to distinct points $x_1$ and $x_2$ respectively. Then from
      the Hausdorff property of $\tilde{X}$, there are disjoint open sets
      $\mathcal{O}_1$ and $\mathcal{O}_2$ containing $x_1$ and $x_2$
      respectively. From continuity of $\pi$, the pre-images of
      $\mathcal{O}_1$ and $\mathcal{O}_2$ are open. The pre-images contain
      $x_1$ and $x_2$ respectively, and by construction, must be disjoint.
      Hence $\tilde{X}$ is also Hausdorff. In the second case,
      $\tilde{x_1}$ and $\tilde{x_2}$ are mapped to the same image $x\in
      X$. Since $\pi$ is a covering space, there must exist an open set
      $\mathcal{O}$ containing $x$ such that the pre-image of $\mathcal{O}$
      is a union of disjoint open sets each homeomorphic to $\mathcal{O}$
      by the restriction of $\pi$. Then $\tilde{x_1}$ and $\tilde{x_2}$
      must be contained in distinct open sets of the pre-image, since $\pi$
      restricted to each open set is a bijection. So we have shown that in
      all cases, $\tilde{X}$ is Hausdorff. \\

      Now we show that $\tilde{X}$ is compact. Let
      $\tilde{\mathcal{C}}=\{\mathcal{U}_i\}_{i\in I}$ be an infinite open
      cover of $\tilde{X}$. Using $\tilde{\mathcal{C}}$, we build an open
      cover $\mathcal{C}$ of $X$ - for each $x\in X$, we form an open
      subset $\mathcal{O}$ containing $x$ following the steps below:
      \begin{itemize}
        \item Let $\mathcal{O}'$ be the open cover containing $x$ given by
          the definition of the covering space property of $\pi$.
        \item Then $\pi^{-1}(\mathcal{O}')$ is a disjoint union of a finite
          number of open sets open sets $U_1,\ldots,U_k$ that are
          homeomorphic to $\mathcal{O}'$ under the restriction map of
          $\pi$.
        \item For each $1\leq i\leq k$, let $U_i''$ be the open set in
          $\tilde{\mathcal{C}}$ that contains $\pi^{-1}(x)$, and let
          $O_i=\pi(U_i''\cap\pi^{-1}(\mathcal{O}'))$.
        \item Finally, let $\mathcal{O}=O_1\cap\ldots\cap O_k$, and put
          $\mathcal{O}$ into $\mathcal{C}$.
        \item By construction, $\mathcal{O}$ contains $x$, is open, and
          $\pi^{-1}(\mathcal{O})$ is a disjoint union of open sets
          homeomorphic to $\mathcal{O}$ under the restriction of $\pi$ such
          that each disjoint open set is contained in some $\mathcal{U}_i$.
      \end{itemize}
      Now $\mathcal{C}$ is an open cover of $X$, so by compactness of $X$,
      has a finite subcover $\mathcal{O}_1,\ldots,\mathcal{O}_n$. By
      construction, $\pi^{-1}(\mathcal{O}_i)$ is contained in a finite
      number of covers in $\tilde{\mathcal{U}}$. Taking the union of all
      these covers for $1\leq i\leq n$ will give us a finite cover of
      $\mathcal{C}$, by surjectivity of $\pi$. Hence, $\tilde{X}$ is
      compact. \\

      Next we prove the converse of the claim. Assume that $\tilde{X}$ is
      compact Hausdorff. We first show that $X$ is compact. Let
      $\{\mathcal{O}_i\}_{i\in I}$ be an infinite open cover of $X$. By
      continuity and surjectivity of $\pi$,
      $\{\pi^{-1}(\mathcal{O}_i)\}_{i\in I}$ will be an open cover of
      $\tilde{X}$. Then by compactness of $\tilde{X}$, there exists a
      finite set of covers
      $\pi^{-1}(\mathcal{O}_1),\ldots,\pi^{-1}(\mathcal{O}_k)$ that covers
      $\tilde{X}$. Then by surjectivity of $\pi$, the finite set
      $\mathcal{O}_1,\ldots,\mathcal{O}_k$ will cover $X$, implying that
      $X$ is compact. \\

      Finally, we show that $X$ is Hausdorff. Let $x_1$, $x_2$ be two
      distinct points in $X$. Let $\mathcal{O}_1$, $\mathcal{O}_2$ be open
      sets containing $x_1$ and $x_2$ respectively given by the definition
      of the covering space property of $\pi$. For $i=1,2$, let
      $\mathcal{U}_i\subseteq\pi^{-1}(\mathcal{O}_i)$ be one of the open
      sets of $\tilde{X}$ that is homeomorphic to $\mathcal{O}_i$ under the
      restriction map of $\pi$. Let $\tilde{x_i}\in\mathcal{U}_i$ be the
      pre-image of $x_i$. Then $\tilde{x_1}$ and $\tilde{x_2}$ are
      distinct, so by Hausdorff property of $\tilde{X}$, they must be
      contained in disjoint open sets $\mathcal{U}_1'$ and $\mathcal{U}_2'$
      respectively. For $i=1,2$, let
      $\mathcal{O}_i'=\pi(\mathcal{U}_i\cap\mathcal{U}_i')$. Then since the
      restriction of $\pi$ to $\mathcal{U}_i$ is a homeomorphism,
      $\mathcal{O}_i'$ is an open set containing $x_i$. Also, from
      disjointness of $\mathcal{U}_0'$ and $\mathcal{U}_1'$,
      $\mathcal{O}_0'$ and $\mathcal{O}_1'$ must also be disjoint, which
      completes our proof that $X$ is Hausdorff.
    \end{proof}

  \item Let $X$ be a Hausdorff space and $G$ be a group acting on $X$.
    Assume the following two conditions hold.
    \begin{itemize}
      \item The action is \textit{free}, i.e. the stabilizer of every point
        in $X$ is trivial.
      \item The action is \textit{properly discontinuous}, i.e. for all
        $x\in X$, there exists a neighborhood $U$ of $x$ such that the set
        $\{g\in G\,|\, g(U)\cap U\neq\emptyset\}$ is finite.
    \end{itemize}
    Prove that the action of $G$ on $X$ is a covering space action.

    \begin{proof}
      Given $x\in X$, let $U$ be the open neighbourhood of $x$ given by the
      second condition, and let $g_1,\ldots,g_k\in G$ be the distinct
      non-identity elements of $G$ such that $g_i(U)\cap U\neq\emptyset$.
      Starting at $U_0=U$, we iteratively shrink the neighborhood
      containing $x$ $k$-times to get $V=U_k$, and such that $g_i(V)\cap
      V=\emptyset$ for $1\leq i\leq k$. Then this neighborhood $V$ would be
      the witness of $G$ being a covering space action. \\

      Given $U_i$, to obtain $U_{i+1}\subseteq U_i$, let $A$ and $B$ be
      disjoint neighborhoods of $x$ and $g_{i+1}\cdot x$ respectively. Note
      that because the action is free, $g_{i+1}\cdot x\neq x$, and then
      because $X$ is Hausdorff, neighborhoods $A$ and $B$ exist. Set
      $U_{i+i}$ to $U_i\cap A\cap g_{i+1}^{-1}(B)$. Then $U_{i+1}$ would
      be a neighborhood of $x$, and is disjoint from $g_{i+1}(U_{i+1})$
      because $g_{i+1}(U_{i+1})\subseteq B$, which is disjoint from
      $A\supseteq U_{i+1}$, as we wish to show.
    \end{proof}

  \item Let $\pi:\tilde{X}\rightarrow X$ be a covering space and let
    $f:Y\rightarrow X$ be an arbitrary continuous map. Define
    \begin{equation*}
      f^*(Y) = \{(y,\tilde{x})\in Y\times\tilde{X}\,|\,
      f(y)=\pi(\tilde{x})\subset Y\times\tilde{X}\}
    \end{equation*}
    and let
    \begin{equation*}
      f^*(\pi):f^*(Y)\rightarrow Y
    \end{equation*}
    be the restriction of the projection $Y\times\tilde{X}\rightarrow Y$ to
    the first factor. Prove that $f^*(\pi):f^*(Y)\rightarrow Y$ is a
    covering map.

    \begin{proof}
      Let $y\in Y$. Consider the open set $O=f^{-1}(V)$, where
      $V$ is the neighborhood of $f(y)\in X$ that witnesses the covering
      space property of $\pi$. We show that $O$ is a neighborhood of $y$
      that witnesses the covering map property of $f^*(\pi)$. \\

      By choice of $O$, $O$ is a neighborhood of $y$. Let
      $\mathcal{U}=\{U_i\}_{i\in I}$ be all the disjoint sets of open
      subsets of $\tilde{X}$ that are homeomorphic to $V$ under the
      restriction map $\pi$. Let
      $O_i=\{(y,\tilde{x}):\pi(\tilde{x})=f(y),\tilde{x}\in U_i\}\subseteq
      Y\times\tilde{X}$. Then chasing definitions, each $O_i$ is disjoint,
      the pre-image of $O$ under $f^*(\pi)$ is the union of the $O_i$'s,
      and the restriction of $f^*(\pi)$ on each $O_i$ is a bijection. Also,
      each $O_i$ is open in $Y\times\tilde{X}$ because it is the
      intersection of the open subset $Y\times U_i$ of $Y\times\tilde{X}$.
      The restriction of $f^*(\pi)$ on $O_i$ is continuous since it is a
      projection map, which is continuous. It remains to show that the
      restriction of $f^*(\pi)$ on $O_i$ is open, so that the restriction
      of $f^*(\pi)$ on each $O_i$ is a homeomorphism. We show the map is
      open by definitions: It suffices to show that given open sets $A\in
      O$ and $B\in U_i$, the image of $(A\times B)\cap O_i$ under
      $f^*(\pi)$ is open in $Y$. Chasing definitions, this image will be
      $A\cap f^{-1}(\pi(B))$, which will be open in $Y$. This completes our
      proof.
    \end{proof}

  \item Set $X=\mathbb{R}^2\setminus\{0\}$. Define an action of the
    additive group $\mathbb{Z}$ on $X$ via the formula
    \begin{align*}
      n\cdot(x,y) = (2^nx,2^{-n}y)  &&  (n\in\mathbb{Z},(x,y)\in X).
    \end{align*}

    \begin{enumerate}
      \item Prove that this is a covering space action.
        \begin{proof}
          We first verify that this is a group action. Clearly, if $n=0$,
          then the action is the identity. Also, given $m,n\in\mathbb{Z}$,
          it is straightforward to verify that
          \begin{align*}
            n\cdot(m\cdot(x,y)) &= n\cdot(2^mx,2^{-m}y) \\
                                &= (n+m)\cdot(x,y). \\
          \end{align*}
          Finally, for a fixed $n\in\mathbb{Z}$, the map on $X$ involves
          only linear functions, hence $\mathbb{Z}$ acts on $X$
          continuously. Therefore, the formula provided is a group action.
          \\

          We now show that the formula provided gives a covering space
          action. Fix $(x,y)\in X$. Consider the open neighborhood
          $U=(x-\delta,x+\delta)\times\mathbb{R}$ in $X$ for some
          $0<\delta<|x|$. We want to choose a small enough $\delta$ such
          that $n\cdot U$ does not intersect $U$ for all
          $n\in\mathbb{Z}-\{0\}$. Given any $n$, $U$ gets mapped to
          $n\cdot U=(2^nx-2^n\delta,2^nx+2^n\delta)\times\mathbb{R}$. In
          order for $U$ to not intersect $n\cdot U$, it suffices to pick
          $\delta$ such that $|2^nx-2^n\delta|>|x+\delta|$ for all
          $n\in\mathbb{Z}^+$, and $|2^nx+2^n\delta|<|x-\delta|$ for all
          $n\in\mathbb{Z}^-$ . Rearranging the formula, we get
          \begin{equation*}
            \delta < \left|1-\frac{2}{2^n+1}\right||x|,
          \end{equation*}
          for all $n\in\mathbb{Z}-\{0\}$, which will be satisfied by
          setting $\delta<|x|/8$. Hence, given $(x,y)\in X$, the open
          neighborhood $U=(x-|x|/8,x+|x|/8)\times\mathbb{R}$ containing
          $(x,y)$ will witness the covering space action of the group.
        \end{proof}

      \item Prove that the quotient $X/\mathbb{Z}$ is not Hausdorff.
        \begin{proof}
          Denote the quotient map from $X$ to $X/\mathbb{Z}$ by $\pi$.
          Consider the points $p_1=[(0,1)]$ and $p_2=[(1,0)]$ in
          $X/\mathbb{Z}$. Note that these are distinct points because the
          $x$-coordinate of the elements in $\pi^{-1}(p_1)$ is always 0,
          while the $x$-coordinate of the elements in $\pi^{-1}(p_2)$ is
          always non-zero. We show that we cannot separate these two points
          with disjoint open sets. Given an open set $U_i$ containing
          $p_i$, its pre-image of the quotient map must be open in $X$. We
          can assume without loss of generality that pre-images of the open
          sets contain a ball of radius $\epsilon$ centered on $(0,1)$ and
          $(1,0)$ respectively. Let $k\in\mathbb{Z}^+$ be such that
          $2^{-k}<\epsilon$. Then point $[(2^{-k},1)]$ is contained in
          $U_1$ and point $[(1,2^{-k})]$ is contained in $U_2$.
          However, these these two points are the same in the quotient
          space because $k\cdot(2^{-k},1)=(1,2^{-k})$, implying that $U_1$
          and $U_2$ can never be disjoint.
        \end{proof}

      \item Explain how $X/\mathbb{Z}$ is the union of four subspaces
        homeomorphic to $S^1\times\mathbb{R}$ coming from the complementary
        components of the $x$-axis and the $y$-axis.

        \begin{proof}
          Consider the four subsets of $X$ given by
          $(0,\infty)\times\mathbb{R}$, $(-\infty,0)\times\mathbb{R}$,
          $\mathbb{R}\times(0,\infty)$, and $\mathbb{R}\times(-\infty,0)$.
          Each of these subsets is closed under group action, and the four
          subsets form an open cover of $X$. We show that each of these
          subsets is homeomorphic to $S^1\times\mathbb{R}$. Consider one of
          the subsets $(0,\infty)\times\mathbb{R}$. The group action
          identifies each line $x=r$ to the line $x=2r$, which makes this
          subset homeomorphic to $S^1\times\mathbb{R}$. A similar argument
          can be made for each of the other three subsets. Hence,
          $X/\mathbb{Z}$ is the union of four subspaces homeomorphic to
          $S^1\times\mathbb{R}$.
        \end{proof}
    \end{enumerate}
\end{enumerate}
\end{document}
