\documentclass{article}[12pt]
\usepackage{url}
\usepackage{fullpage}

\begin{document}
\begin{center}{\bf MATH 10350: Calculus A Tutorial} \end{center}
{\bf TA:} Liling\\
{\bf Office:} B26 Hayes-Healy\\
{\bf Office Hours:} Monday 6:00-7:00pm, Thursday 3:00-4:00pm, or by appointment\\
{\bf E-mail:} lko@nd.edu\\
{\bf Course Website:} \url{https://www3.nd.edu/~m10350/}\\

\noindent {\bf Objective:} This tutorial will review the more difficult
concepts from MATH 10350 in a smaller group setting. We will develop
problem-solving skills and practice exam-type questions to better prepare
you for the exams.\\

\noindent {\bf Course grade:} The tutorial is worth a total of 25 points
towards your overall course grade (600 points total). Your tutorial grade
is broken up into the following components: 60\% worksheets, 20\%
attendance and participation, 20\% quizzes. Each group should submit one
worksheet each week. Worksheets are due in my mailbox in Hurley 256 by noon
on Friday. Late worksheets will count as a zero. The lowest two worksheet
and attendance scores will be dropped. \\

\noindent {\bf Attendance:} There is no make up for tutorials. In the case
of serious illness, family emergencies or an absence because of a
university sponsored event, you must obtain documentation from the First
Year of Studies office. Please contact your first year advisor in this
case. Absences due to athletic events will be excused only if they are
pre-approved by the university and you have a written note from the
athletic department. Unexcused absences are zeros. If you leave before
being dismissed, it will be marked as an unexcused absence. \\

\noindent {\bf Office hours:} Students are encouraged to attend office
hours regularly for help with homework problems or to review concepts and
material discussed in class. When you come speak with me, you should:
\begin{itemize}
  \item Put in a good attempt at the problem yourself first. 
  \item Bring along any work you've done while attempting the problem. 
\end{itemize}
{\bf Other resources:} Students can attend the office hours of any
instructor or TA for this course. These will be posted under {\bf
Instructors' Info} on the course webpage. The {\bf Learning Resources
Center} can also provide you with assistance, tutoring and group study
sessions. Additionally, free one-on-one tutoring is available in the
Mathematics Library. Students may sign-up at:

\centerline{\url{http://library.nd.edu/mathematics/}}
\noindent For additional information on these resources, please refer to
the Course Information Packet handed out on the first day of class. \\

\noindent{\bf The Honor Code:} This class is conducted under the Notre
Dame Honor Code. Violations of the honor code such as cheating and copying
are not tolerated. Exams and quizzes are closed book and taken alone. We do
not allow calculators on exams or quizzes.
\end{document}
