\subsection{Stability Problem}

\begin{goal}[Mostowski, 1996]
  There are many formulae $F$ strongly representing $Z$ (the relation
  $\text{PROOF}_T(x,y)$) in $T$; Godel's theorem is valid only for one such
  formulae. It is not immediately obvious why the theorem proved for this
  formula
  \[T \not\vdash (\forall x)\; [\neg\text{PROOF}_T(x, \ulcorner
  0=1\urcorner)]\]

  should have philosophical importance while a similar theorem obtained by
  a different choice of formula strongly representing $Z$ (could be) simply
  false.
\end{goal}

\begin{notation}
  $T$ is any theory like PA or $Q$.
\end{notation}

\begin{definition}[CON(T)]
  This is the formula
  \[(\exists x)\; [\text{FORM}_T(x) \wedge \neg\text{Pr}_T(x)].\]

  Note that $\text{Pr}_T(x)$ is c.e., and $\text{FORM}_T(x)$ is computable.
\end{definition}

\begin{definition}[CON'(T)]
  This is the formula
  \[(\forall x)\; [\neg\text{PROOF}_T(x, \ulcorner 0=1\urcorner)].\]

  Note that $\text{PROOF}_T$ is computable.
\end{definition}

\begin{definition}[Strongly Representable]
  A relation $R$ (on some set $S \subseteq$ universe of $\mathcal{M}$, a
  model of $T$) is said to be strongly representable with respect to the
  theory if there is a formula $\phi(\bar{x})$ such that 

  \begin{enumerate}
    \item $\bar{a}\in R \implies T\vdash \phi(\bar{a})$,
    \item $\bar{a}\not\in R \implies T\vdash \neg\phi(\bar{a})$.
  \end{enumerate}
\end{definition}

\begin{fact}[Strongly Representable]
  The strongly representable relations are precisely the computable ones
  (Cooper's chapter on Godel).
\end{fact}

\begin{definition}[Weakly Representable]
  A relation $R$ is such that there is a formula $\phi(\bar{x})$ such that

  \begin{enumerate}
    \item $\bar{a}\in R \implies T\vdash \phi(\bar{a})$,
    \item $\bar{a}\not\in R \implies T\not\vdash \phi(\bar{a})$.
  \end{enumerate}
\end{definition}

\begin{fact}[Weakly Representable]
  The weakly representable relations are precisely the c.e. relations.
\end{fact}

\begin{fact}
  The arithmetic sets are strongly representable over PA.
\end{fact}

\begin{definition}[G2(T)]
  This is the formula
  \[T \neg\vdash CON(T).\]
\end{definition}

\begin{fact}
  $PA \not\vdash \ulcorner G2(PA)\urcorner$.
\end{fact}

\begin{fact}
  $PA + CON(PA) \vdash \ulcorner G2(PA)\urcorner$.
\end{fact}

\begin{fact}
  $TA \vdash \ulcorner G2(PA)\urcorner$.
\end{fact}

\begin{fact}[Godel 1931]
  For any sentence $W$ such that
  \[PA+W \vdash \ulcorner G2(PA)\urcorner,\]
  we have $PA \not\vdash \ulcorner W\urcorner$.
\end{fact}

\begin{goal}[Conditions $\mathcal{C}$]
  One must isolate a set of conditions $\mathcal{C}$ (Hilbert-Bernays-Lob
  permeability conditions) on formulae of $T$ such that every formula of
  $T$ that can reasonably be taken to express the consistency of $T$
  satisfies $\mathcal{C}$, and the following definition (GEN-G2) holds:
\end{goal}

\begin{definition}[GEN-G2]
  If $T$ is consistent and $CON(T)$ is a formula satisfying conditions
  $\mathcal{C}$, then $CON(T)$ is not provable in $T$.
\end{definition}

\begin{goal}[Operator]
  Can we find an operator satisfying a property like ``$T$ is consistent''?
  Want set of conditions $\mathcal{C}$ which takes as input $T$, and
  outputs a sentence like $CON(T)$. Give conditions on the operator.
\end{goal}

\begin{fact}
  For any $\phi(x,y)$ that strongly represents the same set as
  $\text{PROOF}_T(x,y)$,
  \[T \not\vdash \forall x\; \neg\phi(x, \ulcorner 0=1\urcorner).\]
\end{fact}

\begin{question}
  Suppose there exists a TM $\Gamma$ such that $\Gamma(t)$ gives a sentence
  $\phi^t$ true in $\omega$ iff $T$ is consistent and spits $\Gamma(t)$
  spit out $\phi^t$ false in $\omega$ otherwise. If there exists such a TM
  (or OM) spitting out $\phi^t$ true in $\omega$, say $\phi^t$ represents
  ``$T$ is consistent'', how might one come up with a sentence $W$ which

  \begin{enumerate}
    \item ``Adequately expresses the consistency of $T$''
    \item Possible that $P \vdash W$?
  \end{enumerate}

  If $PA+W \not\vdash G1(PA)$, then Godel's proof of G2 wouldn't hold for
  this $W$.
\end{question}

\begin{question}
  For all sentences $\phi^t$ computably representing ``$T$ is consistent'',
  is it the case that
  \begin{enumerate}
    \item $T+\phi^t \not\vdash G1(T)$?
    \item Is it every possible that $T \vdash \phi^t$?
  \end{enumerate}
\end{question}

\begin{question}
  Suppose that $T$ is finitely axiomatizable so that $\ulcorner T\urcorner$
  makes sense. Consider a relation $R(x)$ such that
  \begin{enumerate}
    \item $R(x)$ holds if $X=\ulcorner T\urcorner$ such that $T$ is
      consistent
    \item $\neg R(x)$ otherwise
  \end{enumerate}
\end{question}
