\begin{frame}{WKL $\nleq_{\text{soc}}$ WWKL}
  \begin{itemize}
    \item Fix instance of WKL with only one path $A\in[\omega]^\omega$, and
      where the class of paths computing $A$ has 0 measure.
    \item Given arbitrary instance of WWKL, since set of its paths has
      positive measure, some path must fail to compute $A$.
    \item Any non-recursive $A$ works, as we show next.
      $\blacksquare$
  \end{itemize}
\end{frame}

\begin{frame}{Class of sets computing non-recursive is null}
  \textbf{Proof idea:} Assume not. Then there are enough paths computing
  $A$ such that we can effectively ask them to ``vote'' on when an element
  is in $A$. Thus computing $A$, $\Rightarrow\Leftarrow$.

  \vspace{1em}
  Fix Turing functional $\Phi$. Write $\mathcal{B} :=\{X:
  \Phi^X=A\}$. Note $\mathcal{B}$ is Borel and thus measurable.
  Since the measure of a countable union of null sets is 0, and
  there are only countably many Turing functionals, it suffices to show
  $\mu(\mathcal{B})=0$. Assume $\mu(\mathcal{B})=4m>0$.
\end{frame}

\begin{frame}{Class of sets computing non-recursive is null}
  \begin{itemize}
    \item Approximate $\mathcal{B}$ by open cover
      $\mathcal{O}\supseteq\mathcal{B}$ with
      $\mu(\mathcal{O}-\mathcal{B})<m$.
    \item Approximate $\mathcal{O}$ by basic open sets
      $[\sigma_0],\ldots,[\sigma_n] \subseteq\mathcal{O}$ with
      \[\mu(\mathcal{O}-([\sigma_0]\cup\ldots\cup[\sigma_n])) <m.\]
    \item Fix effective enumeration of branches $\sigma$ in
      $[\sigma_0]\cup\ldots\cup[\sigma_n]$.
    \item To decide if $n\in A$, wait till a measure of $\geq2m$ of such
      $\sigma$'s agree, then use their vote.
    \item Since $[\sigma_0]\cup\ldots\cup[\sigma_n]$ approximate
      $\mathcal{B}$ tightly, such $\sigma$'s exist, and they all lie in
      $\mathcal{B}$. Thus
      \begin{align*}
        \mu(\text{``true'' voters})>2m,\\
        \mu(\text{``false'' voters})<2m. &\;\blacksquare
      \end{align*}
  \end{itemize}
\end{frame}

\begin{frame}{Recursive-Encodability, $\Pi_1^0$-Encodability}
  It remains to prove WKL, WWKL $\nleq_{\text{soc}}$ RT.

  \begin{define}[Recursive-Encodable $A\subseteq\omega$]
    $A\subseteq\omega$ is recursively-encodable if given any
    $X\in[\omega]^\omega$, some subset of $X$ computes $A$.
  \end{define}

  \begin{define}[$\Pi_1^0$-encodable $\mathcal{A}\subseteq\omega^\omega$]
    $\mathcal{A}\subseteq \omega^{\omega}$ is $\Pi_1^0$-encodable if for
    every $X\in[\omega]^\omega$, there exists $Y\subseteq X$ and tree
    $T\leq_T Y$ such that
    \[\mathcal{A} \supseteq [T]\neq\emptyset,\]
    and some $f\leq_T Y$ bounds every branch in $T$.
  \end{define}
\end{frame}

\begin{frame}{Characterizing Encodability}
  \begin{main-thm*}
    $\mathcal{A}\subseteq\omega^{\omega}$ compact, contains no
    $\Sigma_1^1$-subset ($\neq\emptyset$). Then there exists infinite set
    $X$ whose subsets compute no path in $\mathcal{A}$.
  \end{main-thm*}

  \begin{coro*}[Characterizing Recursively-Encodable $A\subseteq\omega$]
    $A\subseteq\omega$ recursively-encodable $\Rightarrow$ $A$ is
    hyperarithmetic.
  \end{coro*}

  \vspace{1em}
  From Main Theorem, suffices to show
  \begin{itemize}
    \item $A\subseteq\omega$ is hyperarithmetic $\Leftrightarrow$
      $\{A\}\subseteq\omega^\omega$ is $\Sigma_1^1$
    \item $A\subseteq\omega$ recursively-encodable $\Leftrightarrow$
      $\{A\}\subseteq\omega^\omega$ is $\Pi_1^0$-encodable
  \end{itemize}
\end{frame}

\begin{frame}{$\{A\}$ is $\Sigma_1^1$ $\Rightarrow$ $A$ hyperarithmetic}
  \begin{align*}
    &\{A\}\subseteq\omega^\omega\; \text{is}\; \Sigma_1^1\\
    \Rightarrow\; & \{A\}= \{g:(\exists f)(\forall s)\; [Q(f\restriction
      s,g\restriction s)]\}\\
    \Rightarrow\; &
      \begin{cases}
        A=\{n:(\exists g)\; [g(n)=1\; \wedge\; (\exists f)(\forall s)\;
          Q(f\restriction s,g\restriction s)]\}\\
        A=\{n:(\forall g)\; [g(n)=0\; \rightarrow\; (\forall f)(\exists s)\;
          \neg Q(f\restriction s,g\restriction s)]\}\\
      \end{cases}\\
    \Rightarrow\; &A\in\Delta_1^1\\
    \Leftrightarrow\; &A\subseteq\omega\; \text{hyperarithmetic}.\;
    \blacksquare\\
  \end{align*}
\end{frame}

\begin{frame}{$A$ recursively-encodable $\Rightarrow$
$\{A\}$ is $\Pi_1^0$-encodable}
  \begin{align*}
    \;&A\subseteq\omega\; \text{recursively-encodable}\\
    \Leftrightarrow\; & (\forall X\in[\omega]^\omega)(\exists Y\subseteq
      X)\; [Y\geq_T A]\\
    \Rightarrow\; & (\forall X\in[\omega]^\omega)(\exists Y\subseteq X)\;
      [\{A\}=\{\Phi^Y\}]\\
    \Rightarrow\; &\{A\}\subseteq\omega^\omega\; \text{is}\;
      \Pi_1^0\text{-encodable}.\; \blacksquare\\
  \end{align*}
\end{frame}

\begin{frame}{WKL $\nleq_{\text{soc}}$ RT}
  \begin{itemize}
    \item Fix instance of WKL with only one path $f$, and that path is
      not hyperarithmetic.
    \item Fix instance $c$ of RT, assume by contradiction all
      its solutions compute $f$.
    \item Given arbitrary $X\in[\omega]^\omega$, relativize $c$ to
      $X$ to get $c$-homogeneous $Y\in[X]^\omega$.
    \item By assumption, $Y$ computes $f$. Since $X$ is arbitrary, $f$ is
      recursively-encodable.
    \item But recursive-encodability is the same as being
      hyperarithmetic, $\Rightarrow\Leftarrow$. $\blacksquare$
  \end{itemize}
\end{frame}

\begin{frame}{$A$ hyperarithmetic $\Rightarrow$ $\{A\}$ is $\Sigma_1^1$}
  \begin{align*}
    \;&A\subseteq\omega\; \text{hyperarithmetic}\\
    \Rightarrow\; & A\in\Sigma_1^1\\
    \Rightarrow\; & A=\{n:(\exists f)(\forall s)\; [R(f\restriction s,n)]\}\\
    \Rightarrow\; & \{A\}= \{g:(\exists f)(\forall s)(\forall n)\\
    &[R(f\restriction s,n) \rightarrow g(n)=1 \wedge \neg R(f\restriction s,n)
      \rightarrow g(n)=0]\}\\
    \Rightarrow\; &\{A\}\subseteq\omega^\omega\; \text{is}\; \Sigma_1^1.\;
    \blacksquare\\
  \end{align*}
\end{frame}

\begin{frame}{$A$ recursively-encodable $\Leftarrow$
$\{A\}$ is $\Pi_1^0$-encodable}
  Assume
  \[(\forall X\in[\omega]^\omega)(\exists Y\subseteq X) [\{A\}=[R^Y]].\]
  Given $A\restriction n$, to decide if $n\in A$, enumerate all nodes
  $\sigma\in2^{<\omega}$ extending $A\restriction n$ and check if
  $R^Y(\sigma)$. Since $A$ is the unique path in $[R^Y]$, by weak Konig's
  lemma, eventually exactly one of $[(A\restriction n)^\frown 0]$ or
  $[(A\restriction n)^\frown 1]$ will be covered by a finite set of basic
  open covers $[\sigma_0]\cup\ldots\cup[\sigma_m]$ with $\neg R(\sigma_i)$.
  Then the other $[(A\restriction n)^\frown j]$ gives the correct value
  of $A(n)$.
  \vspace{0.5em}

  Thus $Y\geq_TA$, making $A$ recursively-encodable. $\blacksquare$
\end{frame}

\begin{frame}{$\mathcal{A}$ contains $\Sigma_1^1$-subset
$\Rightarrow$ $\mathcal{A}$ $\Pi_1^0$-encodable}
  \begin{itemize}
    \item Fix $\mathcal{A}$ containing $\mathcal{B} =\{X: (\exists
      Y)(\forall s)\; R(X,Y,s)\} \neq\emptyset$.
    \item Fix $X\in\mathcal{B}$, $Y$ such that $(\forall s)\;R(X,Y,s)$.
    \item Define $f(n)=\max(X(n),Y(n))$.
    \item Then $f$ is a $\Pi_1^0$-modulus of $\mathcal{A}$, i.e. $f$ is
      dominating enough that every $g\geq f$ computes a $g$-bounded tree
      $T$ whose paths all lie in $\mathcal{A}$:
    \item Given $g\geq f$, the $\Pi_1^{0,g}$-set of paths
      \[\{X\leq g:(\forall s)(\exists \sigma\in 2^s, \sigma\leq g)(\forall
      s'<s)\; R(X,\sigma,s')\}\]
      is bounded by $g$, contains $X$, and is contained in $\mathcal{A}$.
    \item Given arbitrary $X\in[\omega]^\omega$, remove enough elements to
      make a sparse $Y\in[X]^\omega$ with $p_Y>f$.
    \item Then $p_Y$ witnesses $\Pi_1^0$-encodability of $\mathcal{A}$ for
      $X$. $\blacksquare$
  \end{itemize}
\end{frame}

\begin{frame}{Possible additional slides}
  \begin{itemize}
    \item Prove Gandy-Harrington topology (relativized) is Baire
    \item Prove every $\Sigma_1^1$-subset of $2^\omega$ contains an
      $O$-recursive
    \item Prove Galvin-Prikry for open sets
  \end{itemize}
\end{frame}

\begin{frame}{Gandy-Harrington topology is Baire}
  \begin{thm*}[Gandy-Harrington topology is Baire]
    The topology on $\omega^\omega$ where open sets are generated by the
    $\Sigma_1^{1}$-sets is known as the Gandy-Harrington topology. This
    topology is Baire, i.e. the countable intersection of dense
    $\Sigma^1_1$- sets is dense.
  \end{thm*}

  \vspace{1em}
  \textbf{Pf:} Given dense $\Sigma^1_1$-sets
  $\mathcal{D}_0,\mathcal{D}_1,\ldots$ and $\Sigma^1_1$-set
  $\mathcal{U}=\{f:(\exists g)\; R(f,g)\}$, iteratively construct
  $f\in\bigcap_{n\in\omega}\mathcal{D}_n$ and $g$ witnessing
  $f\in\mathcal{U}$:

  \vspace{1em}
  At stage $n$, $[f\restriction n]\cap\mathcal{U} \neq\emptyset$, witnessed
  by an extension of $g\restriction n$. Now the class of reals in
  $[f\restriction n]$ that lie in $\mathcal{U}$ by a witness extending
  $g\restriction n$ is a $\Sigma^1_1$-set, so this class intersects
  $\mathcal{D}_0\cap\ldots\cap\mathcal{D}_{n+1}$. Thus one can choose
  $f\restriction (n+1) \succ f\restriction n$ in this intersection, and
  $g\restriction (n+1) \succ g\restriction n$ witnessing $[f\restriction
  (n+1)]\cap\mathcal{U} \neq\emptyset$. $\blacksquare$
\end{frame}

