\begin{frame}{Notations}
  \begin{itemize}
    \item $\omega=\mathbb{N}$, $a,b,\ldots\in\omega$,
      $A,B,\ldots\subseteq\omega$, $\sigma,\tau\in\omega^{<\omega}$,
      $\mathcal{A},\mathcal{B},\ldots\subseteq\omega^\omega$.
    \item Trees $T$ are subsets of $\omega^{<\omega}$ closed under initial
      segments; i.e. $\sigma^{\frown}n\in T \rightarrow \sigma\in T$.
    \item Paths are functions $f:\omega\rightarrow\omega$. A tree contains
      a path $f$ iff $T$ contains every initial segment $f\restriction n
      \prec f$.
    \item $[T] :=\{f:f\; \text{is a path in}\; T\}$, $[\sigma]:=\{f:f\;
      \sigma\prec f\}$.
    \item Turing functionals on trees are partial-recursive functionals
      $\Gamma:2^{<\omega}\rightarrow\omega^{<\omega}$ where
      $\tau\in\Gamma^{\sigma^\frown n} \rightarrow
      \tau\in\Gamma^{\sigma}$. $\Gamma$ can be extended naturally to
      $\Gamma:2^\omega\rightarrow\omega^{\omega}$.
    \item Given set $A$ and cardinal $\kappa$, $[A]^\kappa
      :=\{B\subseteq A: |B|=\kappa\}$.
    \item Given $A\in[\omega]^\omega$, we sometimes refer to $A$ by its
      characteristic function $c_A:\omega\rightarrow\{0,1\}$, or by its
      principal function $p_A:\omega\rightarrow\omega$, which is the
      strictly increasing function listing the elements of $A$.
  \end{itemize}
\end{frame}

\begin{frame}{Ramsey's Theorem ($\text{RT}$)}
  \begin{define*}[$c$-homogeneous]
    Given $k$-coloring $c:[\omega]^n\rightarrow k$, a subset
    $A\subseteq\omega$ is $c$-homogeneous if every $n$-tuple over $A$ is
    given the same color by $c$.
  \end{define*}

  \vspace{1em}
  \begin{thm*}[Ramsey's Theorem]
    Fix $n,k\leq1$. $\text{RT}_k^n$ is the statement that every $k$-coloring
    $c:[\omega]^n\rightarrow k$ has an infinite $c$-homogeneous set.\\
    RT is the statement $(\forall n)(\forall k)\; \text{RT}_k^n$.
  \end{thm*}

  \vspace{1em}
  RT asserts homogeneity exists.
\end{frame}

\begin{frame}{Randomness}
  Intuitively, a path $f\in2^\omega$ is random if it cannot be ``caught''
  by any descending enumeration of trees $2^{<\omega}\supseteq T_0\supseteq
  T_1\supseteq \ldots$, because such trees would reveal information on the
  sequences caught by all of them. A descending sequence is eligible to judge
  for randomness only if the trees shrink fast enough for the class of paths
  caught by all of them to be ``sparse'', because if too many paths are
  caught, the information on the paths is not useful.

  \vspace{1em}
  One way of formalizing the notion of a sparseness of a tree is:
  \begin{define*}
    A (measurable) binary tree $T\subseteq2^{<\omega}$ has positive measure
    if
    \[\lim_s \frac{|\{\sigma\in T: |\sigma|=s\}|}{2^s} >0.\]
  \end{define*}
\end{frame}

\begin{frame}{WWKL asserts Randomness}
  A series of trees $T_0\supseteq T_1\supseteq\ldots$ is eligible to
  test for randomness if the limit of their measures is 0. Since there are
  only countably many Turing machines that can give randomness tests,
  and the class of non-randoms from each test has measure 0, and the
  measure of a countable union of (measurable) classes of measure 0 is
  still 0, the class of non-randoms has measure 0.

  \vspace{2em}
  Thus if a (measurable) tree has positive measure, it must contain a
  random. Weak Weak Konig's Lemma (WWKL) asserts that at least one of its
  paths is random by stating
  \begin{thm*}[Weak Weak Konig's Lemma]
    Every binary tree with positive measure contains a path.
  \end{thm*}
\end{frame}

\begin{frame}{WWKL, WKL, KL}
  The paths of an instance of Weak Konig's Lemma (WKL), however, may
  all be non-randoms, because WKL includes also binary trees with measure
  0:
  \begin{thm*}[Weak Konig's Lemma]
    Every binary tree with infinite nodes contains a path.
  \end{thm*}

  \vspace{2em}
  Konig's Lemma includes even trees that are not binary:
  \begin{thm*}[Konig's Lemma]
    Every finitely-branching tree with infinite nodes contains a path.
  \end{thm*}
\end{frame}

\begin{frame}{Homogeneity versus Randomness}
  A natural question to ask is which of the statements RT, KL, WKL, WWKL is
  stronger. Intuitively, WKL is stronger than WWKL, because a binary tree
  of positive measure has infinite nodes, so if all binary trees with
  infinite nodes have a path, then those with positive measure must
  have one too. A question that is harder to answer is:
  \newtheorem*{question*}{Question}
  \begin{question*}
    Which notion is stronger, homogeneity or randomness?
  \end{question*}
  If a given a path is homogeneous, does that say anything about how random
  it is? Likewise, if a given path is random, will it necessarily provide
  information on finding homogeneous paths?

  \vspace{0.5em}
  Since the notions of homogeneity and randomness are captured by RT and
  WWKL respectively, we are essentially asking which of RT and WWKL is
  stronger.
\end{frame}

\begin{frame}{Problem, Instance, Solution}
  We formalize the notion of problems and the strength between them.

  \vspace{1em}
  \begin{define*}[Problem, instance, solution]
    A mathematical \textbf{problem} is a collection of \textbf{instances},
    with a collection of \textbf{solutions} for each instance.
  \end{define*}

  \vspace{1em}
  For example, RT is a problem; its instances are the collections of
  colorings $c:[\omega]^n\rightarrow k$ for every $n,k\in\omega$, and the
  solutions of each $c$ are the class of $c$-homogeneous paths.

  \vspace{1em}
  Similarly, WWKL is a problem; its instances are the collections of binary
  trees with positive measure, and the solutions of each instance is the
  class of paths in the tree.
\end{frame}

\begin{frame}{Strongly Omnisciently Computably Reducible
($\leq_{\text{soc}}$)}
  \begin{define*}[Strongly Omnisciently Computably Reducible]
    Problem \textbf{P} is \textit{soc}-reducible to problem \textbf{Q}
    (written \textbf{P} $\leq_{\text{soc}}$ \textbf{Q}) if for every
    \textbf{P}-instance \textit{I}, there is a \textbf{Q}-instance
    \textit{J} such that every solution to \textit{J} computes a solution
    to \textit{I}.
  \end{define*}

  $\leq_{\text{soc}}$ successfully captures this expected relation:
  \begin{center}
    WWKL $\leq_{\text{soc}}$ WKL $\leq_{\text{soc}}$ KL.
  \end{center}

  \vspace{0.5em}
  \textbf{WWKL $\leq_{\text{soc}}$ WKL:} Given a binary tree with
  positive measure, this tree has infinite branches and is therefore an
  instance of WKL. The desired relation follows from definition of
  $\leq_{\text{soc}}$.

  \vspace{0.5em}
  \textbf{WKL $\leq_{\text{soc}}$ KL:} Given a binary tree with infinite
  branches, this tree is finitely-branching and therefore an instance of
  WKL. The desired relation follows from definition of $\leq_{\text{soc}}$.
\end{frame}

\begin{frame}{KL, RT $\leq_{\text{soc}}$ WKL}
  WKL is powerful under $\leq_{\text{soc}}$. Given any KL or RT
  instance, fix one of its solutions, and choose the WKL instance whose only
  path codes that solution as an path in $2^\omega$. By definition of
  $\leq_{\text{soc}}$, the given instance reduces to this WKL instance.

  \vspace{1em}
  For example, given a KL instance, fix one of its solutions
  $f:\omega\rightarrow\omega$, code this solution as a path in $2^\omega$
  in a recoverable manner.

  \vspace{1em}
  For instance, if $f$ is $2,9,3,\ldots$, we can code $f$ as the path
  $110$ $1111111110$ $1110\ldots$.

  \vspace{1em}
  Similarly, any solution of an RT instance can be seen as an infinite
  subset of $\omega$. Its characteristic function will be a path in
  $2^\omega$.
\end{frame}

\begin{frame}{Homogeneity ``independent from'' Randomness}
  Dependencies under $\leq_{\text{soc}}$
  \vspace{2em}

  \begin{center}
    \begin{tikzpicture}[node distance=4cm,auto,thick,>=latex']
      \node (KL) {KL$\leftrightarrow$WKL};
      \node (WWKL) [below right of=KL] {WWKL};
      \node (RT) [below left of=KL] {RT};
      \draw[->] (KL) -- (RT);
      \draw[->] (KL) -- (WWKL);
      %\draw [->,red] (RT) -- coordinate (m) (WWKL);
      %\draw[shift={(m)},red](-0.1,-0.1)--(0.1,+0.1);
    \end{tikzpicture}
  \end{center}
\end{frame}
