\begin{frame}{Galvin-Prikry}
  \begin{fact*}[Galvin-Prikry]
    Given infinite set $Z\subseteq\omega$, node $\sigma\in\omega^{<\omega}$,
    and Turing functional $\Phi:2^\omega\rightarrow \omega^{<\omega}$,
    there exists an infinite subset $X\subseteq Z$ where one of these
    holds:\\

    \vspace{1em}
    \textbf{Positive-solution:} All inf. subsets of $X$ compute (via
    $\Phi$) $\sigma$ \\
    \textbf{Negative-solution:} All inf. subsets of $X$ don't compute (via
    $\Phi$) $\sigma$
  \end{fact*}

  \vspace{1em}
  \textbf{Pf (sketch):} The class of subsets of $Z$ that compute (via
  $\Phi$) $\sigma$ can be coded to form an open set in $2^\omega$, by the
  finite use principle. Being open provides enough ``structure'' for
  regions of homogeneity to exist. $\square$
\end{frame}

\begin{frame}{Set whose subsets cannot compute reals in $\mathcal{C}$}
  \newtheorem*{main-lemma*}{Main Lemma}
  \begin{main-lemma*}[Monin, Patey, 2016]
    \begin{align*}
      \text{Given } &\begin{cases}
        \mathcal{C}\subseteq\omega^\omega \text{ compact, contains
        no nonempty } \Sigma^{1,Z}_1\text{-set},\\
        \text{Turing functional } \Phi:2^{\omega}\rightarrow
        \omega^{<\omega},
      \end{cases}
    \end{align*}
    $Z$ contains an infinite subset $X\subseteq Z$ whose subsets cannot
    compute (via $\Phi$) any real in $\mathcal{C}$.
  \end{main-lemma*}

  \vspace{3em}
  \textbf{Pf:} \textbf{Case 0:} There is a node outside
  $\mathcal{C}$ with a positive GP-solution $X$. This $X$ works since
  all its subsets compute (via $\Phi$) a node outside $\mathcal{C}$.
\end{frame}

\begin{frame}{Set whose subsets cannot compute reals in $\mathcal{C}$
(cont.)}
  \textbf{Case 1:} Every node outside $\mathcal{C}$ has only negative
  GP-solutions, and $\mathcal{C}$ has arbitrarily long nodes with
  positive GP-solutions. By compactness of $\mathcal{C}$, this set of
  nodes contains a real in $\mathcal{C}$. Then
  \begin{align*}
    \emptyset\neq\; &\{f:(\forall \sigma\prec f)\; [\sigma\; \text{has
    positive GP-solutions}]\}\\
    \subseteq\; &\{f:(\forall \sigma\prec f) (\exists \text{ inf }
    X\subseteq Z) (\forall \text{ fin } X_0\subset X) (\exists \text{ fin }
    X_1\subset X)\\ &\;\;\;\;\;\;[X_0 \prec X_1 \wedge
    \Phi^{X_1}(\sigma)\downarrow]\}
  \end{align*}
  is a nonempty $\Sigma_1^{1,Z}$-set contained
  in $\mathcal{C}$, $\Rightarrow\Leftarrow$.
\end{frame}

\begin{frame}{Set whose subsets cannot compute reals in $\mathcal{C}$
(cont.)}
  \textbf{Case 2:} For some $n$, every node $\sigma_0,\ldots,\sigma_m$ in
  $\mathcal{C}$ of length $n$ has only negative GP-solutions. Iterate GP
  across them to get decreasing subsets of negative solutions
  \[Z=X_0 \supseteq X_1 \supseteq \ldots\supseteq X_m=X,\]
  where $X_{i+1}\subseteq X_i$ is a negative GP-solution under inputs
  $X_i$ and $\sigma_i$. $X_{i+1}$ exists since GP gives only negative
  solutions for each $\sigma_i$ under $Z$, giving also only negative
  solutions under $X_i\subseteq Z$.

  \vspace{0.5em}
  Take $X=X_m$. Then every subset of $X$ cannot compute (via $\Phi$) any
  node in $\mathcal{C}$ of length exceeding $n$, and in particular,
  cannot compute any real in $\mathcal{C}$. $\blacksquare$

  \vspace{0.5em}
  \textbf{Note:} Each $X_{i+1}$ is picked from the $\Sigma^{1,X_i}_1$-set
  \begin{align*}
    &\{Y:Y\subseteq X_i \text{ inf } \wedge Y \text{ is a
    negative GP-solution for } \sigma_i]\}\\
    =\;&\{Y:Y\subseteq X_i \text{ inf } \wedge (\forall \text{ fin }
    Y_0\subset Y)\; [\Phi^{Y_0}(\sigma_i)\uparrow]\}.
  \end{align*}
\end{frame}

\begin{frame}{Set whose subsets cannot compute reals in $\mathcal{C}$
(relativized)}
  \begin{main-lemma*}[\textcolor{red}{Relativized}]
    \begin{align*}
      \text{Given } &\begin{cases}
        \mathcal{C}\subseteq\omega^\omega \text{ compact, contains
        no nonempty } \Sigma^{1,Z}_1\text{-set},\\
        \text{Turing functional } \Phi:2^{\omega}\rightarrow
        \omega^{<\omega},
      \end{cases}
    \end{align*}
    $Z$ contains an infinite subset $X\subseteq Z$ whose subsets cannot
    compute (via $\Phi$) any real in $\mathcal{C}$.\\
    \vspace{0.5em}
    \textcolor{red}{Furthermore, we can choose $X$ such that $\mathcal{C}$
    also contains no nonempty $\Sigma_1^{1,X}$-subset.}
  \end{main-lemma*}

  \textbf{Pf:} In the proof of the non-relativized lemma, $Z=X_0
  \supseteq \ldots\supseteq X_m=X$, and $X_{i+1}$ is picked from a
  $\Sigma^{1,X_i}_1$-set. The immunity-basis theorem asserts
  that every such set contains some $X_{i+1}$ such that $\mathcal{C}$
  contains no $\Sigma_1^{1,X_{i+1}}$-set, if $\mathcal{C}$ contains no
  $\Sigma^{1,X_i}_1$-subset. Iterate this theorem across $X_i$'s.
  $\blacksquare$
\end{frame}

\begin{frame}{Preserving initial segment}
  \begin{main-lemma*}[Relativized, \textcolor{red}{Segment-preserving}]
    \begin{align*}
      \text{Given } &\begin{cases}
        \mathcal{C}\subseteq\omega^\omega \text{ compact, contains
        no nonempty } \Sigma^{1,Z}_1\text{-set},\\
        \text{Turing functional } \Phi:2^{\omega}\rightarrow
        \omega^{<\omega},
      \end{cases}
    \end{align*}
    $Z$ contains an infinite subset $X\subseteq Z$ whose subsets cannot
    compute (via $\Phi$) any real in $\mathcal{C}$.\\
    \vspace{0.5em}
    Furthermore, we can choose $X$ such that $\mathcal{C}$ also contains no
    nonempty $\Sigma_1^{1,X}$-subset.\\
    \vspace{0.5em}
    \textcolor{red}{Finally, given $s\in\omega$, $X$ can be chosen to
    preserve $Z\restriction s$.}
  \end{main-lemma*}

  \vspace{1em}
  \textbf{Pf:} The idea is to iterate the relativized lemma across all
  subsets of $Z\restriction s$. Iteration is possible after relativization.
\end{frame}

\begin{frame}{Preserving initial segment (cont.)}
  List the subsets of $Z\restriction s$ as $d_0,\ldots,d_n$. Iterate the
  relativized lemma across them to get decreasing subsets
  \[Z=X_0 \supseteq X_1 \supseteq X_2 \supseteq\ldots \supseteq X_n.\]
  At stage $i$, apply lemma with inputs $X_i$, and $\Phi_i$ defined by
  \[\Phi_i(Y) =\Phi(Y\cup d_i),\]

  giving $X_{i+1}\subseteq X_i$ whose subsets cannot compute (via
  $\Phi_i$) reals of $\mathcal{C}$, and where $\mathcal{C}$ contains no
  nonempty $\Sigma_1^{1,X_i}$-set.

  \vspace{1em}
  Choose $X=X_n\cup Z\restriction s$. Since $X$ almost equals $X_n$ and
  $\mathcal{C}$ contains no $\Sigma_1^{1,X_n}$-set, $\mathcal{C}$ also
  contains no $\Sigma_1^{1,X}$-set. Let $Y\subseteq X$ with $Y\restriction
  i=d_i$. Stage $i$ ensured that the subsets of $Y$ compute no real in
  $\mathcal{C}$ via $\Phi_i$. Since $Y\supset d_i$, the same is ensured via
  $\Phi$. $\blacksquare$
\end{frame}

\begin{frame}{Set whose subsets cannot compute reals in $\mathcal{C}$}
  \begin{main-thm*}[Monin, Patey, 2016]
    $\mathcal{C}\subseteq\omega^\omega$ compact, contains no
    $\Sigma_1^1$-subset ($\neq\emptyset$). Then there exists infinite set
    $X$ whose subsets cannot compute any real in $\mathcal{C}$.
  \end{main-thm*}

  \vspace{1em}
  \textbf{Pf:} The main lemma worked for fixed functional $\Phi$. Iterate
  the lemma across all functionals $\Phi_0,\Phi_1,\ldots$ to construct
  decreasing subsets of $X$'s
  \[\omega= X_0\supseteq X_1\supseteq\ldots,\]
  where at stage $s$, the lemma is applied with inputs $X_s$ and
  $\Phi_s$, and $X_{s+1}$ preserves the first $s$-elements of $X_s$. Take
  $X=\bigcap_{s\in\omega}X_s$; this $X$ is infinite from
  segment-preservation. $\blacksquare$
\end{frame}
