\begin{frame}{WKL $\nleq_{\text{soc}}$ WWKL}
  \begin{itemize}
    \item Fix instance of WKL with only one path $A\in[\omega]^\omega$ where
      \[\mu(\{X: X\; \text{computes element in}\; [A]^\omega\}) =0.\]
    \item Given arbitrary instance of WWKL, since set of its paths has
      measure $>0$, some path must fail to compute $A$ 
    \item Any non-computable $A$ works
  \end{itemize}
\end{frame}

\begin{frame}{$A>_T\emptyset \implies\mu(\{X: X\; \text{computes element
in}\; [A]^\omega\}) =0$}
  \textbf{Main idea:} Assume not. Then there are enough paths computing
  $A$ such that we can effectively ask them to ``vote''
  on when an element is in $A$. Thus giving recursive procedure to compute
  $A$, $\Rightarrow\Leftarrow$.
  \vspace{1em}

  Fix Turing functional $\Phi$. Write $\mathcal{B} :=\{X:
  \Phi^X=A\}$. Note $\mathcal{B}$ is Borel and thus measurable.
  Since the measure of a countable union of null sets is 0, and
  there are only countably many Turing functionals, it suffices to show
  $\mu(\mathcal{B})=0$. Assume $\mu(\mathcal{B})=4m>0$.
\end{frame}

\begin{frame}{$A>_T\emptyset \implies\mu(\{X: X\; \text{computes element
in}\; [A]^\omega\}) =0$}
  \begin{itemize}
    \item Approximate $\mathcal{B}$ by open cover
      $\mathcal{O}\supseteq\mathcal{B}$ with
      $\mu(\mathcal{O}-\mathcal{B})<m$.
    \item Approximate $\mathcal{O}$ by basic open sets
      $[\sigma_0],\ldots,[\sigma_n] \subseteq\mathcal{O}$ with
      \[\mu(\mathcal{O}-([\sigma_0]\cup\ldots\cup[\sigma_n])) <m.\]
    \item Fix effective enumeration of branches $\sigma$ in
      $[\sigma_0]\cup\ldots\cup[\sigma_n]$.
    \item To decide if $n\in A$, wait till a measure of $2m$ of such
      $\sigma$'s agree, and take their vote.
    \item Since $[\sigma_0]\cup\ldots\cup[\sigma_n]$ approximate
      $\mathcal{B}$ tightly, such $\sigma$'s exist, and they all lie in
      $\mathcal{B}$. Thus the group of ``true'' voters has measure
      exceeding $2m$.
    \item Also because of the tight approximation, the group of ``false''
      voters has measure less than $2m$.
  \end{itemize}
\end{frame}

\begin{frame}{RT $\nleq_{\text{soc}}$ WWKL}
  \begin{itemize}
    \item $\text{RT}_2^1$ $\nleq_{\text{soc}}$ WWKL
    \item Fix $A\in[\omega]^\omega$ where
      \begin{align*}
        &\mu(\{X: X\; \text{computes element in}\; [A]^\omega\}) =0,\\
        \text{and}\;\;\; &\mu(\{X: X\; \text{computes element in}\;
        [\bar{A}]^\omega\}) =0.\\
      \end{align*}
    \item Choose 2-coloring $c:\omega\rightarrow\{0,1\}$, $c(n)=0
      \Leftrightarrow n\in A$.
    \item Given any instance of WWKL, since the set of its paths has
      measure $>0$, some path must fail to compute any infinite
      homogeneous set of $c$. 
    \item Any bi-hyperimmune set $A$ works.
  \end{itemize}
\end{frame}

\begin{frame}{$A$ Hyperimmune $\implies \mu(\{X: X\; \text{computes
element in}\; [A]^\omega\}) =0$}
  Hyperimmune sets $A$ are characterized as those whose principal
  function $p_A$ is not dominated by any recursive $f$. \\
  \vspace{1em}

  Bi-Hyperimmune sets $A$ are those where $A$, $\bar{A}$ are hyperimmune.\\
  \vspace{1em}

  \textbf{Main idea:} Assume not. Then there are enough paths computing
  infinite subsets of $A$ such that we can effectively ask them to ``vote''
  for when new elements of $A$ have appeared. Thus giving recursive
  $f$ that dominates $p_A$.\\
  \vspace{1em}

  Fix Turing functional $\Phi$. Write $\mathcal{C} :=\{X:
  \Phi^X\in[A]^\omega\}$. Note $\mathcal{C}$ is Borel and thus measurable.
  Since the measure of a countable union of null sets is 0, and
  there are only countably many Turing functionals, it suffices to show
  $\mu(\mathcal{C})=0$. Assume $\mu(\mathcal{C})=4m>0$.
\end{frame}

\begin{frame}{$A$ Hyperimmune $\implies \mu(\{X: X\; \text{computes
element in}\; [A]^\omega\}) =0$ (cont.)}
  \begin{itemize}
    \item Approximate $\mathcal{C}$ by open cover
      $\mathcal{O}\supseteq\mathcal{C}$ with
      $\mu(\mathcal{O}-\mathcal{C})<m$.
    \item Approximate $\mathcal{O}$ by basic open sets
      $[\sigma_0],\ldots,[\sigma_n] \subseteq\mathcal{O}$ with
      \[\mu(\mathcal{O}-([\sigma_0]\cup\ldots\cup[\sigma_n])) <m.\]
    \item Fix effective enumeration of branches $\sigma$ in
      $[\sigma_0]\cup\ldots\cup[\sigma_n]$.
    \item At stage $s=0$, wait till a measure of $2m$ of such $\sigma$'s
      finds an element via $\Phi$; let $f(0)$ be the largest element found.
      Since $[\sigma_0]\cup\ldots\cup[\sigma_n]$ approximate $\mathcal{C}$
      tightly, such $\sigma$'s exist, and more than a measure of $m$ of
      them indeed lie in $\mathcal{C}$.
    \item In particular, at least one of these $\sigma$'s really computes
      an infinite subset of $A$, so $f(0)\geq p_A(0)$.
    \item At stage $s+1$, wait till a measure of $2m$ of $\sigma$'s
      finds an element exceeding $f(s)$; let $f(s+1)$ be the largest
      element found.
  \end{itemize}
\end{frame}

\begin{frame}{Bi-Hyperimmune sets $A$ exist}
  \begin{itemize}
    \item Want $A$ where $p_A,p_{\bar{A}}\not<f$ for all recursive $f$.
    \item Fix $\emptyset''$-enumeration $f_0,f_1,\ldots$ of all recursive
      functions.
    \item At stage $n$, we have defined the first $i$th elements of $A$.
    \item Put enough elements into $\bar{A}$ so that the $(i+1)$th element
      of $A$ exceeds that of $f_n$.
    \item Then we would have defined the first $j$th elements of $\bar{A}$.
    \item Similarly, put enough elements into $A$ so that the $(j+1)$th
      element of $\bar{A}$ exceeds that of $f_n$.
  \end{itemize}
\end{frame}
